### 2023-03-24 LaTeX/innovate.tex
LaTeX update

%\documentclass[%
%reprint,
%%singlecolumn,
%superscriptaddress,
%%groupedaddress,
%%unsortedaddress,
%%runinaddress,
%%frontmatterverbose, 
%%preprint,
%%showpacs,preprintnumbers,
%%nofootinbib,
%%nobibnotes,
%%bibnotes,
%%pre, 
%floatfix,
%amsmath,
%amssymb,
%aps,
%notitlepage
%]{revtex4-1}
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb,graphicx,fullpage}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage[hidelinks=true]{hyperref} 
\usepackage{xcolor}
\usepackage{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm, amsbsy}
\NoCaptionOfAlgo
\usepackage{tcolorbox}

\usepackage[hidelinks=true]{hyperref}
\usepackage{xcolor}
%\definecolor{greenish}{rgb}{0.13,0.58,0.16}
\definecolor{greenish}{RGB}{141,198,63}
\definecolor{reddish}{RGB}{239,65,54}
\definecolor{blueish}{RGB}{28,117,188}
\definecolor{alg1}{RGB}{122,89,239}
\definecolor{alg2}{RGB}{245,147,106}
\definecolor{alg3}{RGB}{124,198,137}

\hypersetup{
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = blueish, %Colour for external hyperlinks
linkcolor    = reddish, %Colour of internal links
citecolor   = greenish %Colour of citations
}

\def\grad{\nabla}
\def\at{\hat{a}}
\def\Ai{\mathcal{A}}
\def\H{\mathcal{H}}
\def\J{\mathcal{J}}
\def\R{\mathbb{R}}
\def\d{\text{d}}
\def\e{\text{e}}
\def\r{\mathbf{r}}
\def\x{\mathbf{x}}
\def\xst{\mathbf{x}^*}
\def\u{\mathbf{u}}
\def\a{\mathbf{a}}
\def\T{\mathsf{T}}
\def\R{\mathbb{R}}
\def\E{\mathbb{E}}
\def\S{\mathcal{S}}
\def\pol{\pi_\theta}
\def\qsa{Q(s, a)}
\def\paa{\pi(a, s)}
\def\H{\sigma}
\def\grad{\nabla}
\def\th{\hat{\mathbf{t}}}
\def\ph{\hat{\mathbf{p}}}
\def\nh{\hat{\mathbf{n}}}
\def\ite{\text{ite}}
\DeclareMathOperator*{\argmax}{arg\,max}

\newtcolorbox{myalgorithm}[2][]{%
    fonttitle=\bfseries,
    title=#2,
    #1,
    before upper={\begin{algorithm}[H]},
    after upper={\end{algorithm}}
}

\begin{document}
\title{Influence of innate behavior on the dynamics of innovation in foraging}
\author{S Ganga Prasath}
%\email{sgangaprasath@iitm.ac.in}
%\affiliation{Department of Applied Mechanics, Indian Institute of Technology Madras, Chennai TN 632406.}
\date{}

%\begin{abstract}
%For later...
%\end{abstract}

%\pacs{Valid PACS appear here}
\maketitle
\section{2D grid world}
Let us say an agent has an intrinsic behavior to follow a herd or a trail
laid by other ants or a flock of birds. All the agents simply following this
herd following behavior will not result in new solutions. Agents often have
to innovate and find new solutions either because the trails are not available
anymore or because you know a better route (influence of history) or perturbations
(intrinsic or environmental) can throw you off trails and you have to find new solutions.
We are interested in understanding the role of intrinsic behavior on the dynamics of innovation.

We will start by implementing an agent trying to optimally traverse from point
A to point B in a 2D grid world. The steps involved in traditional reinforcement
learning based on SARSA algorithm involves essentially 2-steps: $(i)$ action-value
function update, $(ii)$ policy update. The action-value update equation in on-policy SARSA is given by
\begin{align}
Q_\pi(s_t, a_t) \leftarrow & \ Q_\pi(s_t, a_t) \nonumber \\
& \ + \alpha \{ r_{t+1} + \gamma Q_\pi(s_{t+1}, a_{t+1}) - Q_\pi(s_t, a_t) \}
\end{align}
while the policy update is given by
\[
\pi(s) = \argmax_a Q(s, a).
\]
This is shown in algorithmic form in Alg.~\ref{alg:rlsealg}. \\[5pt]

\begin{center}
\begin{tcolorbox}[colframe=alg1!50,colback=alg1!10,fonttitle=\bfseries, title=Algorithm~\ref*{alg:rlsealg}: SARSA algorithm for action-value update]
\begin{algorithm}[H]
\textit{Initialize}:
State, $s_0$;
action, $a_0$;
action-value function, $Q(s_j, a_j)$

\ForEach{epoch}{
Update action-value function using SARSA rule:
    \begin{align*}
    	Q_\pi(s_t, a_t) \leftarrow & \ Q_\pi(s_t, a_t) + \alpha \{ r_{t+1} + \gamma Q_\pi(s_{t+1}, a_{t+1}) - Q_\pi(s_t, a_t) \}
    \end{align*}
 Choose action:
 \begin{align*}
 \pi(s_t) =& \ \argmax_a Q(s_t, a_t)
 \end{align*}
 Calculate reward, $r_t$

 Check if target $s^*$ is reached, continue if not
}
\caption[Recursive least-squares estimator: $\pi \approx \pi^*$]{}
\label{alg:rlsealg}
\end{algorithm}
\end{tcolorbox}
\end{center}

\section{Trail following behavior}
Consider a trail of pheromone $\x^*(s)$ represented by arc-length parameterization $s$.
The concentration field in 2D is then given by $c(\x) = c_o \delta(\x-\x^*(s))$.
This filed of course is assumed to be steady but can be made time-dependent by simply adding a decay time-scale $\tau$ to get: $c(\x, t) = c_o \delta(\x-\x^*(s)) \e^{-t/\tau}$. We can immediately evaluate some of the properties of the curve $\x^*(s)$ which will come in handy soon: $\th(s) = \d \x^*(s)/\d s = \{ \cos \psi(s), \sin \psi(s) \}$ and $\nh(s) = \{ \sin \psi(s), \cos \psi(s) \}$. As we can see the entire curve $\x ^*(s)$ can be represented only using $\psi(s)$ up to global translations and rotations, which is a well known property of curves in 2D.

The state of the agent/ant in our problem is represented by its coordinates
$\r(t)=\{ r_x(t), r_y(t) \}$ and it can make measurements about how far from the
pheromone trail it is $\d\r(t)$ as well as the orientation of the trail, $\th(s, t)$.
The action that it takes from these measurements is to align its orientation, $\ph(t)$
along a direction that will take it towards the trail. We show in Fig.~\ref{fig:schm1}
a schematic of the problem set up.The strategy used by the agent for tracking the pheromone
trail will be to move along the direction given by $(\d \r/|\d \r| + \th)$ by a fixed
length $h$. In order to identify the location along $\xst(s)$ where $\d \r$ intersects,
we use the condition $\d\r(t) \perp \th(s,t)$.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{./figs/schematic.pdf}
    \label{fig:schm1}
    \caption{Schematic of setup}
\end{figure}

\subsection{Semi-circular trail}
For the problem at hand, we will considering a semi-circular trail whose coordinates can be
written in arc-length form as $\xst(s) = \{ a \cos(s/a), a \sin(s/a) \}$ where $a$ is the radius
of the circle. The trail starts at $s=0$ given by coordinates $\xst(0) = \{ a, 0 \}$ and ends at
$s=\pi a$ at $\xst (\pi a) = \{ -a, 0 \}$. From this it is easy to find that $\th(s) = \{ -\sin(s/a) , \cos(s/a) \}
= \{ \cos(\pi/2 + s/a), \sin(\pi/2 + s/a) \}$, and $\nh(s) = \{ -\cos(s/a) , \sin(s/a) \}$. We see that we
can represent $\th(s)$ through $\psi(s) = (\pi/2 + s/a)$. We can now calculate the location along the arc-length
where $\r(t)$ is closest by using the constraint condition $\d\r(t) \perp \th(s,t)$ or equivalently $\d\r(t)
\cdot \th(s,t) = 0$. We can now write $\d \r(t) \sim \{ a \cos(s/a) - r_x, a \sin(s/a) - r_y \}$
(up to normalization) and get the location $s_*(t) = a \tan^{-1}(r_y/r_x)$ by using the formula
for $\th(s)$. From this it is trivial to see that the angle along $\d \r$ is $\phi(t) = \tan^{-1}(r_y/r_x)$
(as is evident in Fig.~\ref{fig:schm1}$(b)$). For a given location of the agent, $\r(t)$ the orientation
it needs to take in the next step can be easily calculated to be $\theta(t) = (\pi/4 - \phi(t))$.

We can now set up the entire dynamics of the agent's trail following behavior on this semi-circle. We can state this
in the notation of reinforcement learning as it will become helpful later. The state of the agent
is $S^t = \{ r_x^t, r_y^t \}$, the measurements it makes are $M^t = \phi^t$ and using this information
the action the agent takes is $A^t = \{ \theta^t \}$ which is its orientation. The dynamics of
the agent can now be written as follows:
\begin{align}
    \text{Measurement update: } \phi^{t+1} = & \ \tan^{-1} \bigg( \frac{r_y^t}{r_x^t} \bigg) + \zeta^t, \\
    \text{Action update: }\ph^{t+1} =& \ \th^{t+1} + \d\r^{t+1}, \\
    \text{State update: } \r^{t+1} =& \ \r^t + l \ph^{t+1},
\end{align}
where $\r^t = \{ r_x^t, r_y^t \}$, $\d\r^t = (\xst(s^t)-\r^t)/||\xst(s^t)-\r(t)||$ and $\ph^t = \{ \cos \theta^t, \sin \theta^t \}$.
We have added sensory noise $\zeta^t$ (which is sampled from a uniform distribution) to the measurement
to reflect the error that accompanies measurements usually. The solution dynamics is shown
in Fig.~\ref{fig:schm1}$(c)$. We call this line-following behavior as a policy
$\pi_\ite(\r, \phi): $
which denotes the agent's innate behavior to follow pheromone trails.

\end{document}

### 2023-04-02 LaTeX/innovate.tex
OUaddition

@@
 \def\nh{\hat{\mathbf{n}}}
+\def\yd{\dot{y}}
 \def\ite{\text{ite}}
@@
 \begin{document}
-\title{Influence of innate behavior on the dynamics of innovation in foraging}
+% \title{Influence of innate behavior on the dynamics of innovation in foraging}
+\title{Dynamics of innovation along foraged trajectories}
 \author{S Ganga Prasath}
@@
 
+\section{Ornstein-Uhlenbeck process}
+
+Let us consider an agent whose orientation $\ph = ( \cos \theta, \sin \theta)$ wants to relax to a
+neutral orientation with stochasticity associated with estimation error. The dynamics of the orientation
+can be written as a stochastic differential equation given by
+\begin{align}
+    \dot{\theta}(t) =& \ - \alpha \theta(t) + \sqrt{2 D} \eta(t), \\
+    \langle \eta(t) \eta(0) \rangle =& \ \delta(t).
+\end{align}
+This is the evolution of the famous Ornstein-Uhlenbeck process which has been extremely well studied
+and the evolution of the probability density can be solved exactly. The Fokker-Plack equation for
+its dynamics which provides the probability of a given state $\theta(t)$ i.e. $p(\theta, t)$, is given by
+\[
+    \frac{\partial p}{\partial t} = \alpha \frac{\partial(p \theta)}{\partial \theta }
+    + D \frac{\partial^2 p}{\partial x^2}
+\]
+The dynamics of the angle relaxes to equilibrium angle $\theta = 0$ in the absence of fluctuations.
+The evolution of the mean is $\dot{h}(t) = - \alpha h(t)$ where $h(t) = \langle \theta(t) \rangle$
+while the co-variance dynamics is given by $\dot{\zeta}(t) = -2\alpha \zeta(t) + 2D$ with
+$\zeta = \langle (\theta(t) - h(t))^2 \rangle = \langle \theta^2(t) \rangle - h(t)^2$.
+
+The position of the agent evolves as $\dot{\r}(t) = v_o \ph(t)$. Under the assumption that the angle
+is small, we can write $\dot{\r}(t) = v_o (1, \theta)$. Thus we have $\yd (t) = v_o \theta$ which can
+be used to rewrite the orientation equation as
+\begin{align}
+    \ddot{y}(t) =& \ - \alpha \yd (t) + v_o\sqrt{2 D} \eta(t).
+\end{align}
+This second order equation needs to boundary condition. We require $y(0) = 0$ and $\yd (0) = 0$ which
+comes from setting the initial orientation to be 0. We are interested in the evolution of the mean and the variance of this SDE. Let $m(t)= \langle y(t) \rangle$
+where the averaging is the ensemble average (equivalently the average of the several trajectories of the
+same dynamics). The mean $m(t)$ evolves as
+\begin{align}
+    \ddot{m}(t) =& \ -\alpha \dot{m}(t).
+\end{align}
+By applying the above boundary condition, we can immediately see that $m(t) = 0$. Nevertheless, we are
+interested in the evolution of the standard deviation, $\beta(t) = \langle [y(t) - m(t)]^2 \rangle$.
+In order to derive an expression for that, we differential the definition once to get
+\begin{align*}
+    \dot{\beta}(t) =& \ 2 \langle \{ y(t) - m(t) \} \{ \yd(t) - \dot{m}(t) \} \rangle, \\
+    = & \ 2 \langle y \yd \rangle - 2 m \dot{m}, \\
+    = & \ 2 v_o [ \langle y \theta \rangle - m \dot{m}].
+\end{align*}
+We can estimate $\langle y \theta \rangle$ as follows
+\begin{align*}
+    \langle y \theta \rangle =& \ v_o \E \bigg[ \int_0^t \theta(s) \theta(t) \ \d s \bigg], \\
+    =& \ v_o  \int_0^t \langle\theta(t) \theta(s)\rangle \ \d s.
+\end{align*}
+We know that for an OU process, 
+\[
+\langle\theta(t) \theta(s)\rangle = \frac{D}{\alpha} \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg].
+\]
+From this we see that
+\begin{align*}
+    \langle y \theta \rangle =& \ \frac{v_o D}{\alpha} \int_0^t  \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg] \ \d s, \\
+    =& \ \frac{v_o D}{\alpha^2} \bigg[ 1 - 2 \e^{-\alpha t} + \e^{-2 \alpha t}\bigg].
+\end{align*}
+We thus have the evolution of the co-variance of $y(t)$ as
+\begin{align}
+    \dot{\beta}(t) = & \  \frac{2 v_o^2 D}{\alpha^2} \bigg[ 1 - 2 \e^{-\alpha t} + \e^{-2 \alpha t}\bigg].
+\end{align}
+We finally have the variance as a function of time as
+\begin{align}
+    \beta(t) & = \frac{v_o^2 D}{\alpha^3} \bigg[ 2 \alpha t - \e^{-2 \alpha t} + 4 \e^{-\alpha t} \bigg].
+\end{align}
+For $t \gg \alpha^{-1}$ we can immediately see that $\beta(t) \approx 2v_o^2 D t/\alpha^2$. We are interested
+in the variance at time, $t^* = c/v_o$ where $c$ is the distance of the agent from the goal A and $v_o$
+is the intrinsic speed of the agent. At this time the location of the agent along x-axis is $x(t^*)=c$
+while the variance in $y(t)$ is $\beta(t^*) \approx 2v_o D c/\alpha^2$.
+% Differentiating this once more, we get
+% \begin{align*}
+%     \ddot{\beta}(t) =& \ 2 [ v_o \langle (\yd \theta + y \dot{\theta} \rangle - \dot{m}^2 - m \ddot{m}]
+% \end{align*}
+
 \end{document}

### 2023-04-03 LaTeX/innovate.tex
correctClear

@@
 \usepackage{bm, amsbsy}
-\NoCaptionOfAlgo
+% `\NoCaptionOfAlgo'
 \usepackage{tcolorbox}
@@
 \def\pol{\pi_\theta}
-\def\qsa{Q(s, a)}
+\def\qsa{Q (s, a)}
 \def\paa{\pi(a, s)}
@@
 }
-\caption[Recursive least-squares estimator: $\pi \approx \pi^*$]{}
-\label{alg:rlsealg}
+\caption[Recursive least-squares estimator: $\pi \approx \pi^*$]{}\label{alg:rlsealg}
 \end{algorithm}
@@
     \centering
-    \includegraphics[width=\textwidth]{./figs/schematic.pdf}
-    \label{fig:schm1}
+    \includegraphics[width=\textwidth]{./figs/schematic.pdf}\label{fig:schm1}
     \caption{Schematic of setup}
@@
 while the co-variance dynamics is given by $\dot{\zeta}(t) = -2\alpha \zeta(t) + 2D$ with
-$\zeta = \langle (\theta(t) - h(t))^2 \rangle = \langle \theta^2(t) \rangle - h(t)^2$.
+$\zeta = \langle {(\theta(t) - h(t))}^2 \rangle = \langle \theta^2(t) \rangle - h{(t)}^2$.
 
@@
 By applying the above boundary condition, we can immediately see that $m(t) = 0$. Nevertheless, we are
-interested in the evolution of the standard deviation, $\beta(t) = \langle [y(t) - m(t)]^2 \rangle$.
+interested in the evolution of the standard deviation, $\beta(t) = \langle {[y(t) - m(t)]}^2 \rangle$.
 In order to derive an expression for that, we differential the definition once to get

### 2023-04-19 LaTeX/innovate.tex
updates
Different formalism

@@
 \[
-    \frac{\partial p}{\partial t} = \alpha \frac{\partial(p \theta)}{\partial \theta }
-    + D \frac{\partial^2 p}{\partial x^2}
+    \frac{\partial p}{\partial t} = \alpha \frac{\partial(\theta p)}{\partial \theta }
+    + D \frac{\partial^2 p}{\partial \theta^2}
 \]
@@
 interested in the evolution of the standard deviation, $\beta(t) = \langle {[y(t) - m(t)]}^2 \rangle$.
-In order to derive an expression for that, we differential the definition once to get
+In order to derive an expression for that, we differentiate the definition once to get
 \begin{align*}
@@
 \begin{align}
-    \beta(t) & = \frac{v_o^2 D}{\alpha^3} \bigg[ 2 \alpha t - \e^{-2 \alpha t} + 4 \e^{-\alpha t} \bigg].
+    \beta(t) & = \frac{v_o^2 D}{\alpha^3} \bigg[ 2 \alpha t - \e^{-2 \alpha t} + 4 \e^{-\alpha t} - 3 \bigg].
 \end{align}
@@
 
+\section{Evaluating value function of policies}
+We are at a stage where we can solve the speed/distance-accuracy trade-off for the an agent trying
+to make optimal decision on when to initiate innovation. In order to do that, let us define clearly
+what is the state and action of the agent.
+\subsection{$v(s^t)$ for $\pi_{\rm int}(s^t)$}
+The agent's state $s^t$ is given by its position, $s^t: \r^t=\{ x^t, y^t \}$ and the action it can take
+at these location is choose a orientation to move, $a^t: \theta^t$. Once the agent choose this action,
+it is taken to the next state $s^t$ through the dynamics: $\r^{t+1} = \r^t + l \ph_\theta^{t}$.
+Before we compute the state-value function for the intrinsic policy, $\pi_{\rm int}(s)$, the sequence
+of process we described above can be written as
+\begin{align}
+    \text{Measurement: } \phi^{t} = & \ \tan^{-1} \bigg( \frac{r_y^t}{r_x^t} \bigg), \\
+    \text{Action, $a^t$: }\theta^{t} =& \ \frac{\pi}{2} + \phi^{t} + \zeta^t, \\
+    \text{State update, $s^{t+1}$: } \r^{t+1} =& \ \r^t + l \ph_\theta^{t}.
+\end{align}
+It is however difficult to evaluate the value function of this specific functional form of the policy.
+Instead we move to a different form given by the Langevin dynamics of the orientation satisfying
+the equation:
+\begin{align}
+    \gamma \partial_j \phi (s, j) =& \ B \partial_s^2 \phi (s, j) + \Gamma(s, j),
+\end{align}
+where the arc-length $s \in (0, L], L = nl$ and epoch $j = 0 \dots N$. The bending stiffness $B$
+in the equation captures the constraint that the motion of the agent is in the forward direction.
+Further this equation is subject to the boundary condition $\phi(0, j) = 0, \phi'(0, j) = \kappa_o$ with $\kappa_o$ being
+the curvature of the initial pheromone trail, $\kappa_o \sim a^{-1}$. Here $\Gamma(s, j)$ is the
+spatio-epochal delta-correlated noise whose variance is given by
+\[
+\langle \Gamma_\alpha \Gamma_\beta \rangle = 2 \gamma D \delta_{\alpha \beta} \delta(s-s') \delta(j-j')
+\]
+
+The interpretation of this equation is the the orientation dynamics of the agent undergoes a diffusion but the overall trajectory has 
+an intrinsic curvature (arising out of the boundary condition). Here $s$ is the arc-length which is
+equivalent to $k$-step motion of the agent i.e. $s_k = k l = k v_o \delta t$. Similarly $j$ in the
+equation is the $j$-th epoch of $k$-step trajectory of the agent.
+
 \end{document}

### 2023-04-19 LaTeX/innovate.tex
SmallUpdate

@@
 equivalent to $k$-step motion of the agent i.e. $s_k = k l = k v_o \delta t$. Similarly $j$ in the
-equation is the $j$-th epoch of $k$-step trajectory of the agent.
+equation is the $j$-th epoch of $k$-step trajectory of the agent. The location of the agent is given by
+$\partial_s \r(s,j) = \th_\theta(s, j) = \{ \cos \phi(s, j), \sin \phi(s, j) \} = \{ -\sin \theta(s, j), \cos \theta(s, j) \}$. Once we know $\phi(s, j)$
+can evaluate the position, $\r(s, j)$ and the orientation $\theta(s, j)$. In this framework
+the policy $\pi_{\rm int} (\r(s))$ for the $j$-th epoch gives the action $\phi(s)$ that moves the agent
+along using the geometric constraint/equation $\r'(s) = \th_\theta(s)$.
+\textbf{Don't need temporal/epochal dynamics}
 

### 2023-04-20 LaTeX/innovate.tex
Checks

--- 2023-04-19 LaTeX/innovate.tex
@@
 \def\ite{\text{ite}}
+\def\P{\mathcal{P}}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 \end{align}
-It is however difficult to evaluate the value function of this specific functional form of the policy.
-Instead we move to a different form given by the Langevin dynamics of the orientation satisfying
-the equation:
-\begin{align}
-    \gamma \partial_j \phi (s, j) =& \ B \partial_s^2 \phi (s, j) + \Gamma(s, j),
-\end{align}
-where the arc-length $s \in (0, L], L = nl$ and epoch $j = 0 \dots N$. The bending stiffness $B$
-in the equation captures the constraint that the motion of the agent is in the forward direction.
-Further this equation is subject to the boundary condition $\phi(0, j) = 0, \phi'(0, j) = \kappa_o$ with $\kappa_o$ being
-the curvature of the initial pheromone trail, $\kappa_o \sim a^{-1}$. Here $\Gamma(s, j)$ is the
-spatio-epochal delta-correlated noise whose variance is given by
+It is however difficult to evaluate the value function for this specific functional form of the policy.
+We assume that the distribution of the orientation of the agent is distributed as through
+it is a worm-like chain polymer whose orientation distribution is given by
 \[
-\langle \Gamma_\alpha \Gamma_\beta \rangle = 2 \gamma D \delta_{\alpha \beta} \delta(s-s') \delta(j-j')
+    \P[\theta(s)] = \frac{\e^{-\beta E_b}}{\mathcal{Z}}
 \]
+where the bending energy $E_b[\theta(s)] = B/2 \int_0^L (\theta'(s)-\kappa_o)^2 \ \d s$, fugacity
+$\mathcal{Z} = \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
+stiffness $B$ captures the constraint for the agent to move forward, $\beta$ is the effective
+temperature capturing the fluctuations in the trajectory of agent. We also have
+boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. We can discretize
+the bending energy as
+\[
+    E_b[\theta] = \frac{B}{2} \sum_0^n l \bigg( \frac{\theta^{n+1} - \theta^n}{l}  - \kappa_o \bigg)^2.
+\]
+Assuming
 
-The interpretation of this equation is the the orientation dynamics of the agent undergoes a diffusion but the overall trajectory has 
-an intrinsic curvature (arising out of the boundary condition). Here $s$ is the arc-length which is
-equivalent to $k$-step motion of the agent i.e. $s_k = k l = k v_o \delta t$. Similarly $j$ in the
-equation is the $j$-th epoch of $k$-step trajectory of the agent.
+% It is however difficult to evaluate the value function of this specific functional form of the policy.
+% Instead we move to a different form given by the Langevin dynamics of the orientation satisfying
+% the equation:
+% \begin{align}
+%     \gamma \partial_j \phi (s, j) =& \ B \partial_s^2 \phi (s, j) + \Gamma(s, j),
+% \end{align}
+% where the arc-length $s \in (0, L], L = nl$ and epoch $j = 0 \dots N$. The bending stiffness $B$
+% in the equation captures the constraint that the motion of the agent is in the forward direction.
+% Further this equation is subject to the boundary condition $\phi(0, j) = 0, \phi'(0, j) = \kappa_o$ with $\kappa_o$ being
+% the curvature of the initial pheromone trail, $\kappa_o \sim a^{-1}$. Here $\Gamma(s, j)$ is the
+% spatio-epochal delta-correlated noise whose variance is given by
+% \[
+% \langle \Gamma_\alpha \Gamma_\beta \rangle = 2 \gamma D \delta_{\alpha \beta} \delta(s-s') \delta(j-j')
+% \]
+
+% The interpretation of this equation is the the orientation dynamics of the agent undergoes a diffusion but the overall trajectory has 
+% an intrinsic curvature (arising out of the boundary condition). Here $s$ is the arc-length which is
+% equivalent to $k$-step motion of the agent i.e. $s_k = k l = k v_o \delta t$. Similarly $j$ in the
+% equation is the $j$-th epoch of $k$-step trajectory of the agent. The location of the agent is given by
+% $\partial_s \r(s,j) = \th_\theta(s, j) = \{ \cos \phi(s, j), \sin \phi(s, j) \} = \{ -\sin \theta(s, j), \cos \theta(s, j) \}$. Once we know $\phi(s, j)$
+% can evaluate the position, $\r(s, j)$ and the orientation $\theta(s, j)$. In this framework
+% the policy $\pi_{\rm int} (\r(s))$ for the $j$-th epoch gives the action $\phi(s)$ that moves the agent
+% along using the geometric constraint/equation $\r'(s) = \th_\theta(s)$.
+% \textbf{Don't need temporal/epochal dynamics}
 

### 2023-04-22 LaTeX/innovate.tex
SmallEdits
Thinking
FunctionFormUpdate

@@
 \def\P{\mathcal{P}}
+\def\D{\mathcal{D}}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 \[
-    \P[\theta(s)] = \frac{\e^{-\beta E_b}}{\mathcal{Z}}
+    \P[\theta(s)] = \frac{1}{\mathcal{Z}}\exp{(-\beta E_b)}
 \]
@@
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. We can discretize
-the bending energy as
+boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. 
+% We can discretize the bending energy as
+% \[
+%     E_b[\theta] = \frac{B}{2} \sum_0^n l \bigg( \frac{\theta^{n+1} - \theta^n}{l}  - \kappa_o \bigg)^2.
+% \]
+We are interested in evaluating the value function of this policy, $v(s^t)$ whose
+expression is given by
 \[
-    E_b[\theta] = \frac{B}{2} \sum_0^n l \bigg( \frac{\theta^{n+1} - \theta^n}{l}  - \kappa_o \bigg)^2.
+    v_\pi(S) = \E_\pi \bigg[ \sum_{k=0}^\infty \gamma^k R_{t+k+1} \bigg| S_t = S \bigg].
 \]
-Assuming
+We can write this in path-integral form as
+\begin{align}
+    v_\pi(s) =& \ \sum_{k=t}^T \int \D[\theta(s)] \P[\theta(s)] \gamma^{k-t} R_{k+1}, \\
+    \approx& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] R(s) \ \d s
+\end{align}
+The reward itself is a function of the agent's azimuthal direction, $\phi(s) = \tan^{-1}(y(s)/x(s))$.
+Further we know the geometric relation $x'(s) = \cos(\theta(s)), y'(s) = \sin(\theta(s))$. The functional
+form of the reward is $R(\phi) = \exp(-\phi/\phi^*)$ and 
+\[
+    \phi(s) = \tan^{-1}\bigg( \frac{\int_0^s \sin(\theta(\tau)) \ \d \tau}{\int_0^s \cos(\theta(\tau)) \ \d \tau} \bigg)
+\]
+We can write the ultimate expression for the value function we would like to evaluate as,
+\begin{align}
+    v_\pi(s) =& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] \exp(-\phi(s)/\phi^*) \ \d s'.
+\end{align}
+We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o$
+which implies we can write $\theta(s) = \kappa_o s + \sum_{n=0}^{n_c} a_n s \sin(n \pi s/L)$.
 

### 2023-04-22 LaTeX/innovate.tex
SmallChng

@@
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. 
+boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$. 
 % We can discretize the bending energy as
@@
 \end{align}
-We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o$
+We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o$.
+Before we solve the stochastic version of the problem, let us solve it for the deterministic case when
+the agent follows the semi-circular path exactly. In this scenario, $\theta(s) = \pi/2 + \kappa_o s$ and
+from this it immediately follows that $\phi(s) = (\pi - \kappa_o s)/2$.
+
 which implies we can write $\theta(s) = \kappa_o s + \sum_{n=0}^{n_c} a_n s \sin(n \pi s/L)$.

### 2023-04-22 LaTeX/innovate.tex
Updates

--- 2023-04-20 LaTeX/innovate.tex
@@
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. We can discretize
-the bending energy as
+boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$. 
+% We can discretize the bending energy as
+% \[
+%     E_b[\theta] = \frac{B}{2} \sum_0^n l \bigg( \frac{\theta^{n+1} - \theta^n}{l}  - \kappa_o \bigg)^2.
+% \]
+We are interested in evaluating the value function of this policy, $v(s^t)$ whose
+expression is given by
 \[
-    E_b[\theta] = \frac{B}{2} \sum_0^n l \bigg( \frac{\theta^{n+1} - \theta^n}{l}  - \kappa_o \bigg)^2.
+    v_\pi(S) = \E_\pi \bigg[ \sum_{k=0}^\infty \gamma^k R_{t+k+1} \bigg| S_t = S \bigg].
 \]
-Assuming
+We can write this in path-integral form as
+\begin{align}
+    v_\pi(s) =& \ \sum_{k=t}^T \int \D[\theta(s)] \P[\theta(s)] \gamma^{k-t} R_{k+1}, \\
+    \approx& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] R(s) \ \d s
+\end{align}
+The reward itself is a function of the agent's azimuthal direction, $\phi(s) = \tan^{-1}(y(s)/x(s))$.
+Further we know the geometric relation $x'(s) = \cos(\theta(s)), y'(s) = \sin(\theta(s))$. The functional
+form of the reward is $R(\phi) = \exp(-\phi/\phi^*)$ and 
+\[
+    \phi(s) = \tan^{-1}\bigg( \frac{\int_0^s \sin(\theta(\tau)) \ \d \tau}{\int_0^s \cos(\theta(\tau)) \ \d \tau} \bigg)
+\]
+We can write the ultimate expression for the value function we would like to evaluate as,
+\begin{align}
+    v_\pi(s) =& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] \exp(-\phi(s)/\phi^*) \ \d s'.
+\end{align}
+We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o$.
+Before we solve the stochastic version of the problem, let us solve it for the deterministic case when
+the agent follows the semi-circular path exactly. In this scenario, $\theta_b(s) = \pi/2 + \kappa_o s$ and
+from this it immediately follows that $\phi_b(s) = \kappa_o s$. For this particular functional form of $\phi(s)$
+we can write the reward as $R(s) = \exp(-\kappa_o s/\phi^*)$. The value function from this can be written as
+\begin{align}
+    v_b(s) =& \ \int_s^L \exp(-\kappa_o \tau/\phi^*) \ \d \tau, \\
+    =& \ -\frac{\phi^*}{\kappa_o} [\e^{-\kappa_o L/\phi^*} - \e^{-\kappa_o s/\phi^*}], \\
+    =& \ \frac{\phi^*}{\kappa_o} \e^{-\kappa_o L/\phi^*} [ \e^{\kappa_o(L-s)/\phi^*} - 1].
+\end{align}
+We now introduce perturbations in the trajectory of the agent from the trajectory $\theta_b(s)$ by introducing
+perturbations $\theta(s) = \theta_b(s) + \delta \theta(s)$.
 

### 2023-04-24 LaTeX/innovate.tex
Update
MoreUpdates

--- 2023-04-22 LaTeX/innovate.tex
@@
 \def\D{\mathcal{D}}
+\def\Z{\mathcal{Z}}
+\def\ah{\hat{a}}
+\def\theta{\vartheta}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 what is the state and action of the agent.
-\subsection{$v(s^t)$ for $\pi_{\rm int}(s^t)$}
+\subsection{$v(\r, \theta)$ for $\pi_{\rm int}(\r, \theta)$}
 The agent's state $s^t$ is given by its position, $s^t: \r^t=\{ x^t, y^t \}$ and the action it can take
@@
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o = a^{-1}$. 
+boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$.
+
 % We can discretize the bending energy as
@@
 \[
-    v_\pi(S) = \E_\pi \bigg[ \sum_{k=0}^\infty \gamma^k R_{t+k+1} \bigg| S_t = S \bigg].
+    v_{\rm int}(\r, \theta) = \E_\pi \bigg[ \sum_{k=0}^\infty \gamma^k R_{t+k+1} \bigg| S_t = S \bigg].
 \]
@@
 \begin{align}
-    v_\pi(s) =& \ \sum_{k=t}^T \int \D[\theta(s)] \P[\theta(s)] \gamma^{k-t} R_{k+1}, \\
+    v_{\rm int}(\r, \theta) =& \ \sum_{k=t}^T \int \D[\theta(s)] \P[\theta(s)] \gamma^{k-t} R_{k+1}, \\
     \approx& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] R(s) \ \d s
@@
 \[
-    \phi(s) = \tan^{-1}\bigg( \frac{\int_0^s \sin(\theta(\tau)) \ \d \tau}{\int_0^s \cos(\theta(\tau)) \ \d \tau} \bigg)
+    \phi(\r, \theta) = \tan^{-1}\bigg( \frac{\int_0^s \sin(\theta(\tau)) \ \d \tau}{\int_0^s \cos(\theta(\tau)) \ \d \tau} \bigg)
 \]
@@
 \begin{align}
-    v_\pi(s) =& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] \exp(-\phi(s)/\phi^*) \ \d s'.
+    v_{\rm int}(\r, \theta) =& \ \int_{s}^L \int \D[\theta(s)] \P[\theta(s)] \exp(-\phi(s)/\phi^*) \ \d s'.
 \end{align}
-We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = 0$ and $\theta'(0) = \kappa_o$
-which implies we can write $\theta(s) = \kappa_o s + \sum_{n=0}^{n_c} a_n s \sin(n \pi s/L)$.
+We know that $\theta(s)$ has to satisfy the boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o$.
+Before we solve the stochastic version of the problem, let us solve it for the deterministic case when
+the agent follows the semi-circular path exactly. In this scenario, $\theta_b(s) = \pi/2 + \kappa_o s$ and
+from this it immediately follows that $\phi_b(s) = \kappa_o s$. For this particular functional form of $\phi(s)$
+we can write the reward as $R(s) = \exp(-\kappa_o s/\phi^*)$. The value function from this can be written as
+\begin{align}
+    v_b(s) =& \ \int_s^L \exp(-\kappa_o \tau/\phi^*) \ \d \tau, \\
+    =& \ -\frac{\phi^*}{\kappa_o} [\e^{-\kappa_o L/\phi^*} - \e^{-\kappa_o s/\phi^*}], \\
+    =& \ \frac{\phi^*}{\kappa_o} \e^{-\kappa_o L/\phi^*} [ \e^{\kappa_o(L-s)/\phi^*} - 1].
+\end{align}
+We now introduce perturbations in the trajectory of the agent from the trajectory $\theta_b(s)$ by introducing
+perturbations $\theta(s) = \theta_b(s) + \delta \theta(s)$. We know that $\phi$ and $\theta$ are related by the
+geometric constraint: $\phi(s) = \theta(s) - \pi/2$ which immediately gives $\phi(s) = \kappa_o s + \delta \theta(s)$.
+We can then write the value function as
+\begin{align}
+    v_{\rm int}(s) =& \ \int_{s}^L \int \frac{ \D[\theta(s)]}{\mathcal{Z}}\e^{-(\beta B/2) \int_0^L (\theta'(s)-\kappa_o)^2 \ \d s} \exp(-\phi(\tau)/\phi^*) \ \d \tau, \\
+    =& \ \int_{s}^L \int \frac{ \D[\theta(s)]}{\mathcal{Z}}\e^{-(\beta B/2) \int_0^L [\delta \theta'(s)]^2 \ \d s} \e^{(-\kappa_o \tau - \delta \theta(\tau))/\phi^*} \ \d \tau.
+\end{align}
+Let us start by calculating $\Z$,
+\begin{align}
+    \Z =& \ \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}, \\
+    =& \ \int \mathcal{D}[\theta(s)] \e^{-(\beta B/2) \int_0^L [\delta \theta'(s)]^2 \ \d s}, \\
+    \delta \theta(s) =& \ \sum_{k=-\infty}^\infty \ah_k \e^{i 2 \pi k s/L}, \\
+    \int_0^L [\delta \theta'(s)]^2 \ \d s =& \ \int_0^L \bigg[ \sum_{k=-\infty}^\infty \frac{2\pi k}{L} \ah_k i \e^{i 2 \pi k s/L} \bigg]
+    \bigg[ \sum_{q=-\infty}^\infty \frac{2\pi q}{L} \ah_q i \e^{i 2 \pi q s/L} \bigg] \ \d s, \\
+    =& \ \frac{4\pi^2}{L} \sum_{k=-\infty}^\infty k^2 \ah_k^2, \text{where we have used } \ah_{-k} = \ah_k \\
+    \Z =& \ \prod_{k=-\infty}^\infty \int_{-\infty}^{\infty} \d \ah_k \e^{-\frac{\beta B k^2}{L} 4\pi^2\ah_k^2}, \\
+    =& \ \prod_{k=-\infty}^\infty \sqrt{\frac{\pi L}{\beta B}} \frac{1}{2\pi k}
+\end{align}
+We can go ahead and write down the value function as
+\begin{align}
+    v_{\rm int}(s) =& \ \frac{1}{\Z} \prod_{k=-\infty}^\infty \int_{-\infty}^{\infty} \e^{-\frac{\beta B k^2}{L} 4\pi^2 \ah_k^2} \bigg[ \int_{s}^L \e^{(-\kappa_o \tau - \ah_k \e^{i 2 \pi k \tau/L})/\phi^*} \ \d \tau \bigg] \d \ah_k, \\
+    =& \ \frac{1}{\Z} \prod_{k=-\infty}^\infty \int_{s}^L \bigg[ \int_{-\infty}^{\infty} \e^{-\frac{\beta B k^2}{L} 4\pi^2 \ah_k^2  - \ah_k \e^{i 2 \pi k \tau/L}/\phi^*} \ \d \ah_k \bigg]  \e^{-\kappa_o \tau/\phi^*} \ \d \tau.
+\end{align}
+We can evaluate the integral inside the square brackets as
+\[
+    \int_{-\infty}^\infty \e^{-\alpha \ah^2 -\gamma \ah} \ \d \ah = \e^{-\frac{\gamma^2}{4\alpha}}\int_{-\infty}^\infty \e^{-\alpha(\ah + \frac{\gamma}{2 \alpha})^2} \ \d \ah
+    = \sqrt{\frac{\pi}{\alpha}} \e^{-\frac{\gamma^2}{4\alpha}}.
+\]
+We then get
+\begin{align}
+    v_{\rm int}(s) =& \ \sqrt{\frac{\pi L}{\beta B}}  \frac{1}{\Z} \prod_{k=-\infty}^\infty \frac{1}{2\pi k}  \int_{s}^L \e^{-\gamma(k, \tau)^2/4\alpha} \e^{-\kappa_o \tau/\phi^*} \ \d \tau.
+\end{align}
+where $\gamma(k, \tau) = \e^{i 2 \pi k \tau/L}/\phi^*, \alpha = \beta B k^2 4 \pi^2/L$. In the limit of small temperature,
+or large bending stiffness, $\alpha \gg 1$ and we can Taylor expand the integral to get
+\begin{align}
+    v_{\rm int}(s) =& \ \sqrt{\frac{\pi L}{\beta B}}  \frac{1}{\Z} \prod_{k=-\infty}^\infty \frac{1}{2\pi k}  \int_{s}^L \bigg[ 1 - \frac{\gamma(k, \tau)^2}{4\alpha} \bigg] \e^{-\kappa_o \tau/\phi^*} \ \d \tau.
+\end{align}
+From this we see that the leading order contribution is still
+\begin{align}
+    v_{\rm int}(s) =& \ \frac{\phi^*}{\kappa_o} \e^{-\kappa_o L/\phi^*} [ \e^{\kappa_o(L-s)/\phi^*} - 1].
+\end{align}
+
+\subsection{$v(\r, \theta)$ for $\pi_{\rm OU}(\r, \theta)$}
+We have seen already the dynamics of the OU process and we can write the policy $\pi_\text{OU}(\r, \theta)$ using the evolution
+of the orientation and position as
+\begin{align}
+    \theta^{t+1} =& \ \theta^t (1 - \alpha \Delta t ) + \sqrt{2 D} \eta^t \Delta t, \\
+    % \dot{\theta}(t) =& \ - \alpha \theta(t) + \sqrt{2 D} \eta(t), \\
+    \langle \eta^t \eta^0 \rangle =& \ \delta(t), \\
+    \r^{t+1} =& \ \r^{t} + l \ph_\theta^t, \text{where} \ \ph_\theta^t = \{ \cos \theta^t, \sin \theta^t \}.
+    % \dot{\r}(t) =& \ v_o \ph
+\end{align}
+We have calculated the evolution of variance of the agent's trajectory from a straight line as a function of
+distance/time from the bifurcation point evolves as
+\begin{align}
+    \psi(t) & = \frac{v_o^2 D}{\alpha^3} \bigg[ 2 \alpha t - \e^{-2 \alpha t} + 4 \e^{-\alpha t} - 3 \bigg].
+\end{align}
+where $\alpha^{-1}$ is the time-scale of relaxation of the agent. The $n$-step variance of the agent is
+$\psi (n \Delta t)$. Assuming that the agent reaching a distance $\sigma$ from the end-point A can reach the
+target, the probability that an agent will reach the target is given by $\P{\rm OU}(A) = \sigma/\psi( n \Delta t)$.
+We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
+\[
+    v_{\rm OU}(\r, \theta) = \int_{s^c}^{s^*} \int \D[\theta(t)] \P[\theta(t)] R(t) \ \d t.
+\]
+Let us assume that the reward function for the OU process is $R(s) = \nu \Theta(||\r(s^*) - \r^*|| - \sigma)$.
 

### 2023-04-26 LaTeX/innovate.tex
SmallChange

@@
 \def\theta{\vartheta}
+\def\H{\mathcal{H}}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 \[
-    \frac{\partial p}{\partial t} = \alpha \frac{\partial(\theta p)}{\partial \theta }
-    + D \frac{\partial^2 p}{\partial \theta^2}
+    \frac{\partial \P}{\partial t} = \alpha \frac{\partial(\theta \P)}{\partial \theta }
+    + D \frac{\partial^2 \P}{\partial \theta^2}
 \]
 The dynamics of the angle relaxes to equilibrium angle $\theta = 0$ in the absence of fluctuations.
-The evolution of the mean is $\dot{h}(t) = - \alpha h(t)$ where $h(t) = \langle \theta(t) \rangle$
+The evolution of the mean is $\dot{h}(t) = - \alpha h(t)$ where $h(t) = \E [ \theta(t) ]$
 while the co-variance dynamics is given by $\dot{\zeta}(t) = -2\alpha \zeta(t) + 2D$ with
-$\zeta = \langle {(\theta(t) - h(t))}^2 \rangle = \langle \theta^2(t) \rangle - h{(t)}^2$.
+$\zeta = \E [  {(\theta(t) - h(t))}^2 ] = \E [  \theta^2(t) ] - h{(t)}^2$.
 
@@
 This second order equation needs to boundary condition. We require $y(0) = 0$ and $\yd (0) = 0$ which
-comes from setting the initial orientation to be 0. We are interested in the evolution of the mean and the variance of this SDE. Let $m(t)= \langle y(t) \rangle$
-where the averaging is the ensemble average (equivalently the average of the several trajectories of the
-same dynamics). The mean $m(t)$ evolves as
+comes from setting the initial orientation to be 0. We are interested in the evolution of the mean and
+the variance of this SDE. Let $m(t)= \E [  y(t) ]$ where the averaging is the ensemble average
+(equivalently the average of the several trajectories of the same dynamics). The mean $m(t)$ evolves as
 \begin{align}
@@
 By applying the above boundary condition, we can immediately see that $m(t) = 0$. Nevertheless, we are
-interested in the evolution of the standard deviation, $\beta(t) = \langle {[y(t) - m(t)]}^2 \rangle$.
+interested in the evolution of the standard deviation, $\beta(t) = \E [  {(y(t) - m(t))}^2 ]$.
 In order to derive an expression for that, we differentiate the definition once to get
 \begin{align*}
-    \dot{\beta}(t) =& \ 2 \langle \{ y(t) - m(t) \} \{ \yd(t) - \dot{m}(t) \} \rangle, \\
-    = & \ 2 \langle y \yd \rangle - 2 m \dot{m}, \\
-    = & \ 2 v_o [ \langle y \theta \rangle - m \dot{m}].
+    \dot{\beta}(t) =& \ 2 \E [ \{ y(t) - m(t) \} \{ \yd(t) - \dot{m}(t) \} ], \\
+    = & \ 2 \E [ y \yd ] - 2 m \dot{m}, \\
+    = & \ 2 v_o \{ \E [ y \theta ] - m \dot{m}\}.
 \end{align*}
-We can estimate $\langle y \theta \rangle$ as follows
+We can estimate $\E [ y \theta ]$ as follows
 \begin{align*}
-    \langle y \theta \rangle =& \ v_o \E \bigg[ \int_0^t \theta(s) \theta(t) \ \d s \bigg], \\
-    =& \ v_o  \int_0^t \langle\theta(t) \theta(s)\rangle \ \d s.
+    \E [ y \theta ] =& \ v_o \E \bigg[ \int_0^t \theta(s) \theta(t) \ \d s \bigg], \\
+    =& \ v_o  \int_0^t \E [ \theta(t) \theta(s) ] \ \d s.
 \end{align*}
@@
 \[
-\langle\theta(t) \theta(s)\rangle = \frac{D}{\alpha} \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg].
+    \E [ \theta(t) \theta(s) ] = \frac{D}{\alpha} \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg].
 \]
@@
 \begin{align*}
-    \langle y \theta \rangle =& \ \frac{v_o D}{\alpha} \int_0^t  \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg] \ \d s, \\
+    \E [ y \theta ] =& \ \frac{v_o D}{\alpha} \int_0^t  \bigg[ \e^{-\alpha(t-s)}-\e^{-\alpha(t+s)} \bigg] \ \d s, \\
     =& \ \frac{v_o D}{\alpha^2} \bigg[ 1 - 2 \e^{-\alpha t} + \e^{-2 \alpha t}\bigg].
@@
 \begin{align}
-    \theta^{t+1} =& \ \theta^t (1 - \alpha \Delta t ) + \sqrt{2 D} \eta^t \Delta t, \\
-    % \dot{\theta}(t) =& \ - \alpha \theta(t) + \sqrt{2 D} \eta(t), \\
-    \langle \eta^t \eta^0 \rangle =& \ \delta(t), \\
+    \theta^{t+1} =& \ \theta^t (1 - \mu \Delta t ) + \sqrt{2 D \Delta t} \zeta^t, \\
+    % \dot{\theta}(t) =& \ - \alpha \theta(t) + \sqrt{2 D} \zeta(t), \\
+    \langle \zeta^t \zeta^0 \rangle =& \ \delta(t), \\
     \r^{t+1} =& \ \r^{t} + l \ph_\theta^t, \text{where} \ \ph_\theta^t = \{ \cos \theta^t, \sin \theta^t \}.
@@
 We have calculated the evolution of variance of the agent's trajectory from a straight line as a function of
-distance/time from the bifurcation point evolves as
+distance/time from the bifurcation point evolves as,
 \begin{align}
-    \psi(t) & = \frac{v_o^2 D}{\alpha^3} \bigg[ 2 \alpha t - \e^{-2 \alpha t} + 4 \e^{-\alpha t} - 3 \bigg].
+    \psi(t) & = \frac{v_o^2 D}{\mu^3} \bigg[ 2 \mu t - \e^{-2 \mu t} + 4 \e^{-\mu t} - 3 \bigg].
 \end{align}
-where $\alpha^{-1}$ is the time-scale of relaxation of the agent. The $n$-step variance of the agent is
-$\psi (n \Delta t)$. Assuming that the agent reaching a distance $\sigma$ from the end-point A can reach the
-target, the probability that an agent will reach the target is given by $\P{\rm OU}(A) = \sigma/\psi( n \Delta t)$.
-We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
+where $\mu^{-1}$ is the time-scale of relaxation of the agent. The $n$-step variance of the agent is
+$\psi (t^*)$, where $t^* = n \Delta t$. Assuming that the agent reaching a distance $\sigma$ from the end-point A can reach the
+target, the probability that an agent will reach the target is given by $\P_{\rm OU}(||\r^{n\Delta t} - \r^*|| \leq \sigma)
+= \sigma/\psi(t^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
 \[
-    v_{\rm OU}(\r, \theta) = \int_{s^c}^{s^*} \int \D[\theta(t)] \P[\theta(t)] R(t) \ \d t.
+    v_{\rm OU}(\r, \theta) = \int_{0}^{t^*} \int \D[\theta(t)] \P_{\rm OU}[\theta(t)] R(t) \ \d t.
 \]
-Let us assume that the reward function for the OU process is $R(s) = \nu \Theta(||\r(s^*) - \r^*|| - \sigma)$.
+Let us assume that the reward function for the OU process is $R(t) = \nu[1 - \Theta(||\r(t^*) - \r^*|| - \sigma)]$.
+From this we see that the value of $v_{\rm OU}(\r^c, \theta^c) = \sigma \nu/\psi(t^*)$.
+
+\section{Value function of combined policy}
+The total value function for the agent's policy is $v(\r, \theta) = v_{\rm int}(\r, \theta) + v_{\rm OU}(\r^c, \theta^c)$.
+The agent is trying to maximize this value function from a given state, $\r^t, \theta^t$. The final expression
+for the value function is
+\[
+    v(\r, \theta) = \frac{\sigma \nu}{\psi(t^*+t_f)} + \frac{\phi^*}{\kappa_o} \e^{-\kappa_o L^*/\phi^*} [ \e^{\kappa_o(L^*-v_ot)/\phi^*} - 1],
+\]
+where $\psi(t) = \dfrac{v_o^2 D}{\mu^3} \bigg[ 2 \mu t - \e^{-2 \mu t} + 4 \e^{-\mu t} - 3 \bigg]$, $t_f$ is the time-taken to reach
+the region around the target location A. In this equation, the diffusion coefficient
+representing the fluctuations in the dynamics of the agent capturing the error in memory
+of target location A. We assume that the fluctuation $D$ is larger farther from the target location.
 

### 2023-04-26 LaTeX/draft.tex
DraftInitiated

--- 2023-04-26 LaTeX/innovate.tex
@@
-%\documentclass[%
-%reprint,
-%%singlecolumn,
-%superscriptaddress,
-%%groupedaddress,
-%%unsortedaddress,
-%%runinaddress,
-%%frontmatterverbose, 
-%%preprint,
-%%showpacs,preprintnumbers,
-%%nofootinbib,
-%%nobibnotes,
-%%bibnotes,
-%%pre, 
-%floatfix,
-%amsmath,
-%amssymb,
-%aps,
-%notitlepage
-%]{revtex4-1}
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb,graphicx,fullpage}
-\usepackage{mathrsfs}
+\documentclass[%
+reprint,
+%singlecolumn,
+superscriptaddress,
+%groupedaddress,
+%unsortedaddress,
+%runinaddress,
+%frontmatterverbose, 
+%preprint,
+%showpacs,preprintnumbers,
+%nofootinbib,
+%nobibnotes,
+%bibnotes,
+%pre, 
+floatfix,
+amsmath,
+amssymb,
+aps,
+notitlepage
+]{revtex4-1}
+%\usepackage[sort&compress,comma,numbers,round]{natbib}
+\usepackage{graphicx}
 \usepackage{bm}
@@
 \usepackage{xcolor}
-\usepackage{algorithm2e}
-\usepackage{mathrsfs}
-\usepackage{bm, amsbsy}
-% `\NoCaptionOfAlgo'
-\usepackage{tcolorbox}
-
 \usepackage[hidelinks=true]{hyperref}
@@
 \definecolor{blueish}{RGB}{28,117,188}
-\definecolor{alg1}{RGB}{122,89,239}
-\definecolor{alg2}{RGB}{245,147,106}
-\definecolor{alg3}{RGB}{124,198,137}
 
@@
 urlcolor     = blueish, %Colour for external hyperlinks
-linkcolor    = reddish, %Colour of internal links
+linkcolor    = magenta, %Colour of internal links
 citecolor   = greenish %Colour of citations
 }
+% \documentclass[12pt]{article}
+% \usepackage{amsmath}
+% \usepackage{amsfonts}
+% \usepackage{amssymb,graphicx,fullpage}
+\usepackage{mathrsfs}
+\usepackage{bm}
+\usepackage[hidelinks=true]{hyperref} 
+\usepackage{xcolor}
+\usepackage{mathrsfs}
+\usepackage{bm, amsbsy}
 
@@
 
-\newtcolorbox{myalgorithm}[2][]{%
-    fonttitle=\bfseries,
-    title=#2,
-    #1,
-    before upper={\begin{algorithm}[H]},
-    after upper={\end{algorithm}}
-}
-
 \begin{document}
-% \title{Influence of innate behavior on the dynamics of innovation in foraging}
-\title{Dynamics of innovation along foraged trajectories}
+\title{Dynamics of innovation in stigmergic collectives}
 \author{S Ganga Prasath}
-%\email{sgangaprasath@iitm.ac.in}
-%\affiliation{Department of Applied Mechanics, Indian Institute of Technology Madras, Chennai TN 632406.}
+\email{sgangaprasath@iitm.ac.in}
+\affiliation{Department of Applied Mechanics, Indian Institute of Technology Madras, Chennai TN 600 036.}
 \date{}
 
-%\begin{abstract}
-%For later...
-%\end{abstract}
+\begin{abstract}
+For later...
+\end{abstract}
 
@@
 \]
-This is shown in algorithmic form in Alg.~\ref{alg:rlsealg}. \\[5pt]
-
-\begin{center}
-\begin{tcolorbox}[colframe=alg1!50,colback=alg1!10,fonttitle=\bfseries, title=Algorithm~\ref*{alg:rlsealg}: SARSA algorithm for action-value update]
-\begin{algorithm}[H]
-\textit{Initialize}:
-State, $s_0$;
-action, $a_0$;
-action-value function, $Q(s_j, a_j)$
-
-\ForEach{epoch}{
-Update action-value function using SARSA rule:
-    \begin{align*}
-    	Q_\pi(s_t, a_t) \leftarrow & \ Q_\pi(s_t, a_t) + \alpha \{ r_{t+1} + \gamma Q_\pi(s_{t+1}, a_{t+1}) - Q_\pi(s_t, a_t) \}
-    \end{align*}
- Choose action:
- \begin{align*}
- \pi(s_t) =& \ \argmax_a Q(s_t, a_t)
- \end{align*}
- Calculate reward, $r_t$
-
- Check if target $s^*$ is reached, continue if not
-}
-\caption[Recursive least-squares estimator: $\pi \approx \pi^*$]{}\label{alg:rlsealg}
-\end{algorithm}
-\end{tcolorbox}
-\end{center}
 
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \caption{Schematic of setup}
-\end{figure}
+\end{figure*}
 
@@
 This is the evolution of the famous Ornstein-Uhlenbeck process which has been extremely well studied
-and the evolution of the probability density can be solved exactly. The Fokker-Plack equation for
+and the evolution of the probability density can be solved exactly. The Fokker-Planck equation for
 its dynamics which provides the probability of a given state $\theta(t)$ i.e. $p(\theta, t)$, is given by
@@
 \[
-    v(\r, \theta) = \frac{\sigma \nu}{\psi(t^*+t_f)} + \frac{\phi^*}{\kappa_o} \e^{-\kappa_o L^*/\phi^*} [ \e^{\kappa_o(L^*-v_ot)/\phi^*} - 1],
+    v(\r, \theta) = \frac{\sigma \nu}{\psi(t^*+t_f)} + \frac{\phi^*}{\kappa_o} \e^{-\kappa_o s^*/\phi^*} [ \e^{\kappa_o(s^*-s)/\phi^*} - 1],
 \]

### 2023-04-27 LaTeX/draft.tex
EditStart

@@
 \maketitle
-\section{2D grid world}
-Let us say an agent has an intrinsic behavior to follow a herd or a trail
-laid by other ants or a flock of birds. All the agents simply following this
-herd following behavior will not result in new solutions. Agents often have
-to innovate and find new solutions either because the trails are not available
-anymore or because you know a better route (influence of history) or perturbations
-(intrinsic or environmental) can throw you off trails and you have to find new solutions.
-We are interested in understanding the role of intrinsic behavior on the dynamics of innovation.
+\section{Problem setting}
+Let us consider a trail of pheromone $\x^*(s)$ representing the position in 2-D coordinates and parameterized by the
+arc-length $s$. The concentration of pheromone is considered continuous along this curve and is given by $c(\x) = c_o \delta(\x-\x^*(s))$.
+This field is of course steady and its dynamics will be considered later in the article. The state of the agent/ant in
+the environment is is represented by its position $\r^t=\{ r_x^t, r_y^t \}$ and the action it can take is to choose the
+direction of motion i.e., its orientation $\ph^t = \{\cos(\theta^t), \sin(\theta^t)\}$, where $[\bullet]^t$ denotes the
+value of the variable at time-instant $t$. The agent can choose the action based on the measurements at each instant of time.
 
-We will start by implementing an agent trying to optimally traverse from point
-A to point B in a 2D grid world. The steps involved in traditional reinforcement
-learning based on SARSA algorithm involves essentially 2-steps: $(i)$ action-value
-function update, $(ii)$ policy update. The action-value update equation in on-policy SARSA is given by
-\begin{align}
-Q_\pi(s_t, a_t) \leftarrow & \ Q_\pi(s_t, a_t) \nonumber \\
-& \ + \alpha \{ r_{t+1} + \gamma Q_\pi(s_{t+1}, a_{t+1}) - Q_\pi(s_t, a_t) \}
-\end{align}
-while the policy update is given by
-\[
-\pi(s) = \argmax_a Q(s, a).
-\]
-
-\section{Trail following behavior}
-Consider a trail of pheromone $\x^*(s)$ represented by arc-length parameterization $s$.
-The concentration field in 2D is then given by $c(\x) = c_o \delta(\x-\x^*(s))$.
-This filed of course is assumed to be steady but can be made time-dependent by simply adding a decay time-scale $\tau$ to get: $c(\x, t) = c_o \delta(\x-\x^*(s)) \e^{-t/\tau}$.
-We can immediately evaluate some of the properties of the curve $\x^*(s)$ which will come in handy soon: $\th(s) = \d \x^*(s)/\d s = \{ \cos \psi(s), \sin \psi(s) \}$ and $\nh(s) = \{ \sin \psi(s), \cos \psi(s) \}$.
-As we can see the entire curve $\x ^*(s)$ can be represented only using $\psi(s)$ up to global translations and rotations, which is a well known property of curves in 2D.
-
-The state of the agent/ant in our problem is represented by its coordinates
-$\r(t)=\{ r_x(t), r_y(t) \}$ and it can make measurements about how far from the
+\subsection*{Trail following behavior}
+Social insects such as ants communicate through pheromone or by antennation. As stigmergy is one of the fundamental
+mechanisms by which the collective coordinates and recruits new workers.
+it can make measurements about how far from the
 pheromone trail it is $\d\r(t)$ as well as the orientation of the trail, $\th(s, t)$.

### 2023-04-27 LaTeX/draft.tex
SomeEdts

@@
 
-\subsection*{Trail following behavior}
-Social insects such as ants communicate through pheromone or by antennation. As stigmergy is one of the fundamental
-mechanisms by which the collective coordinates and recruits new workers.
-it can make measurements about how far from the
-pheromone trail it is $\d\r(t)$ as well as the orientation of the trail, $\th(s, t)$.
-The action that it takes from these measurements is to align its orientation, $\ph(t)$
-along a direction that will take it towards the trail. We show in Fig.~\ref{fig:schm1}
-a schematic of the problem set up.The strategy used by the agent for tracking the pheromone
-trail will be to move along the direction given by $(\d \r/|\d \r| + \th)$ by a fixed
-length $h$. In order to identify the location along $\xst(s)$ where $\d \r$ intersects,
-we use the condition $\d\r(t) \perp \th(s,t)$.
+\subsection*{Trail following behavior for semi-circular trail}
+Social insects such as ants communicate either through pheromone or by antennation.
+As stigmergy is one of the fundamental mechanisms by which the collective coordinates
+and recruits new workers, we assume that the agent can follow the trail $\x^*(s)$.
+This intrinsic behavior is denoted by the policy $\pi_{\rm int}(\r)$ which provides the
+appropriate angle $\theta^t$ required to follow the trail when the agent is at location
+$\r^t$ after the measurment that the agent makes. In order to consider a specific trail
+of the initial pheromone trail, let us assume that the trail is a simple semi-circle with
+functional form $\xst(s) = \{ a \cos(s/a), a \sin(s/a) \}$. For this particular case,
+the policy can be explicitly written as
+\begin{align}
+    \text{Measurement: } \phi^{t} = & \ \tan^{-1} \bigg( \frac{r_y^t}{r_x^t} \bigg) + \zeta^t, \\
+    \text{Action: }\theta^{t} =& \ \frac{\pi}{2} + \phi^{t}, \\
+    \text{State update: } \r^{t+1} =& \ \r^t + l \ph^{t}.
+\end{align}
+where $\zeta^t$ is the fluctuation/error in the measurements that the agent makes, $l$ is the step-length
+of the agent. In the limit when the step-length is small, we can write the continuous version
+of the motion as $\dot{\r} = v_o \ph$. Without loss of generality, we can write this as
+$\d\r/\d s = \ph$ where $s = v_o t$ and we assume that the distribution of the orientation of the agent is distributed as though
+it is a worm-like chain polymer whose orientation distribution is given by
+\[
+    \P[\theta(s)] = \frac{1}{\mathcal{Z}}\exp{(-\beta E_b)}
+\]
+where the bending energy $E_b[\theta(s)] = B/2 \int_0^{s_c} (\theta'(s)-\kappa_o)^2 \ \d s$, fugacity
+$\mathcal{Z} = \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
+stiffness $B$ captures the constraint for the agent to move forward, $\beta$ is the effective
+temperature capturing the fluctuations in the trajectory of agent. We also have
+boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$.
+We are interested in evaluating the value function of this policy $\pi_{\rm int}(\r)$
+denoted by $v_{\rm int}(\r^t)$ whose expression is
+\[
+    v_{\rm int}(\r) = \E_\pi \bigg[ \sum_{k=0}^\infty \gamma^k R_{t+k+1} \bigg| \r_t = \r \bigg].
+\]
+We can write this in path-integral form as
+\begin{align}
+    v_{\rm int}(\r) =& \ \sum_{k=t}^T \int \D[\theta(s)] \P[\theta(s)] \gamma^{k-t} R_{k+1}, \\
+    \approx& \ \int_{s}^{s_c} \int \D[\theta(s)] \P[\theta(s)] R(s) \ \d s
+\end{align}
+The reward itself is a function of the agent's azimuthal direction, $\phi(s) = \tan^{-1}(y(s)/x(s))$.
+Assuming that the reward has the function form $R(\phi) = \exp(-\phi/\phi^*)$, the leading order form of 
+the value function for this policy is
+\[
+    v_{\rm int}(\r) = \frac{\phi^*}{\kappa_o} [ \e^{-\kappa_os/\phi^*} - \e^{-\kappa_o s_c/\phi^*}].
+\]
 

### 2023-04-27 LaTeX/draft.tex
Changes

@@
 value of the variable at time-instant $t$. The agent can choose the action based on the measurements at each instant of time.
+In this article we are interested in understanding how and when an agent can innovate from its innate behavior to follow
+pheromone trail to explore the environment and find novel solutions to the navigation problem.
 
@@
 \]
-where the bending energy $E_b[\theta(s)] = B/2 \int_0^{s_c} (\theta'(s)-\kappa_o)^2 \ \d s$, fugacity
-$\mathcal{Z} = \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
+where the bending energy $E_b[\theta(s)] = B/2 \int_0^{s_c} (\theta'(s)-\kappa_o)^2 \ \d s$, the partition
+function $\mathcal{Z} = \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
 stiffness $B$ captures the constraint for the agent to move forward, $\beta$ is the effective
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$.
-We are interested in evaluating the value function of this policy $\pi_{\rm int}(\r)$
-denoted by $v_{\rm int}(\r^t)$ whose expression is
+boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$. Here $s_c$ is the coordinate
+along the semi-circle at which the agent bifurcates towards the target location A using a new trajectory.
+Our motivation is to identify this location using the Bellman equation which will be discussed later.
+Now in order to evaluate the quality of the intrinsic policy we introduce the notion of value function
+of this policy $\pi_{\rm int}(\r)$,denoted by $v_{\rm int}(\r^t)$. The value function provides the cummulative
+future reward that the agent will acquire when following the policy $\pi_{\rm int}(\r)$. This can be written as
 \[

### 2023-04-27 LaTeX/draft.tex
ModeAddns

@@
 value of the variable at time-instant $t$. The agent can choose the action based on the measurements at each instant of time.
+In this article we are interested in understanding the dynamics of agents that need to innovate in order to find new
+trails over which the collective travels. In order to understand the dynamics, we assume that the agent has already
+traveled along the trajectory $\x^*(s)$ once before.
 In this article we are interested in understanding how and when an agent can innovate from its innate behavior to follow
@@
 \end{figure*}
+
+\subsection*{Search strategy post bifurcation}
+After the agent leaves the trail trajectory at $\r(s_c)$, the agent needs to reach the target A.
 

### 2023-04-27 LaTeX/draft.tex
MoreChngs

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\def\OU{\rm OU}
+\def\intP{\rm int}
 
@@
 In this article we are interested in understanding the dynamics of agents that need to innovate in order to find new
-trails over which the collective travels. In order to understand the dynamics, we assume that the agent has already
-traveled along the trajectory $\x^*(s)$ once before.
-In this article we are interested in understanding how and when an agent can innovate from its innate behavior to follow
-pheromone trail to explore the environment and find novel solutions to the navigation problem.
+trails over which the collective travels.
 
@@
 and recruits new workers, we assume that the agent can follow the trail $\x^*(s)$.
+In this article we are interested in understanding how and when an
+agent can innovate from its innate behavior to follow pheromone trails to explore the environment and find novel solutions
+to the navigation problem. In order to understand the dynamics of innovation, we assume that the agent has already
+traversed the trajectory $\x^*(s)$ before.
 This intrinsic behavior is denoted by the policy $\pi_{\rm int}(\r)$ which provides the
@@
 After the agent leaves the trail trajectory at $\r(s_c)$, the agent needs to reach the target A.
+Though the agent can take a variety of exploration strategies, we assume that the policy resembles
+an Ornstein-Uhlenbeck (OU) process. This explorative policy $\pi_{\OU}(\r)$ can be written through $\P(\theta, t)$
+the probability that the agent will take a particular angle $\theta$ at time $t$ or distance $s$. However since the orientation of the
+agent is dependent on $\r(s_c)$, the angle between agent location $\r(s_c)$ to target A can be denoted
+as $\phi$ and $\theta=\phi+\varphi$ where $\varphi$ is the fluctuation around this orientation that follows
+an OU-process. The probability $\P(\varphi, s)$ can then be written as
+\[
+    \frac{\partial \P}{\partial s} = \frac{\nu}{v_o} \frac{\partial(\varphi \P)}{\partial \varphi }
+    + \frac{D}{v_o} \frac{\partial^2 \P}{\partial \varphi^2}
+\]
+where $\nu$ is the time-scale associated with relaxation of the agent's orientation to the target $\phi$
+and $D$ represents the fluctuation in orientation, which we will soon assume to be large when $\r(s_c)$
+is farther from the target A. We assume that the agent reaching a distance $\sigma$ from the end-point A can reach the
+target and thus the reward is a function of the propagation of variance in agent's position towards the
+target. Instead of solving the partial differential equation for $\P(\varphi, s)$, we can evaluate the
+evolution of the variance of agent's position $\r(s)$ as it marches close to the target.
+
+Assuming that the agent reaching a distance $\sigma$ from the end-point A can reach the
+target, the probability that an agent will reach the target is given by $\P_{\rm OU}(||\r^{n\Delta t} - \r^*|| \leq \sigma)
+= \sigma/\psi(t^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
+\[
+    v_{\rm OU}(\r, \theta) = \int_{0}^{t^*} \int \D[\theta(t)] \P_{\rm OU}[\theta(t)] R(t) \ \d t.
+\]
+Let us assume that the reward function for the OU process is $R(t) = \nu[1 - \Theta(||\r(t^*) - \r^*|| - \sigma)]$.
+From this we see that the value of $v_{\rm OU}(\r^c, \theta^c) = \sigma \nu/\psi(t^*)$.
+
+%\sectionfont{\fontsize{15}{15}\selectfont}
+%\subsectionfont{\fontsize{15}{15}\selectfont}
+%\subsubsectionfont{\fontsize{15}{15}\selectfont}
+\large
+\widetext
+\clearpage
+\onecolumngrid
+\begin{center}
+\textbf{\large Supplemental Information for ``Branching in soft gels''}\\[.2cm]
+S  Ganga  Prasath$^{1, *}$, Aditi Chakrabarti$^{1, *}$,  and  L  Mahadevan$^{1, 2, 3, 4}$\\[.1cm]
+{\small \itshape ${}^1$School of Engineering and Applied Sciences, Harvard University, Cambridge MA 02138.\\
+${}^2$Department of Physics, Harvard University, Cambridge MA 02138.\\
+${}^3$Center  for  Brain  Science,  Harvard  University,  Cambridge,  MA  02138.\\
+${}^4$Department of Organismic and Evolutionary Biology, Harvard University, Cambridge 02138
+}
+\end{center}
+%%%%%%%%% Merge with supplemental materials %%%%%%%%%%
+%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
+\setcounter{equation}{0}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\setcounter{page}{1}
+\makeatletter
+% \renewcommand{\theequation}{S\arabic{equation}}
+% \renewcommand{\thefigure}{S\arabic{figure}}
+% \renewcommand{\bibnumfmt}[1]{[S#1]}
+%\renewcommand{\citenumfont}[1]{S#1}
+%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
+\linespread{1.5}
+% \setstretch{1.5}
+
+%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
+%\renewcommand{\bibnumfmt}[1]{[S#1]}
+%\renewcommand{\citenumfont}[1]{S#1}
+%\renewcommand{\thesection}[1]{S#1\arabic{section}}
+\renewcommand{\thetable}{S\arabic{table}}%
+\renewcommand{\thesection}{S\arabic{section}}
+%\renewcommand{\thesubsection}{SS\arabic{subsection}}
+\renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+%\renewcommand{\thesubsection}{SS\arabic{subsection}{}}
+\renewcommand{\thesubsection}{\roman{subsection}{}}
+%\renewcommand{\thesubsubsection}{(\roman{subsubsection}{})}
 

### 2023-04-28 LaTeX/draft.tex
MinorEdits

@@
 \def\intP{\rm int}
+\def\ebh{\hat{\mathbf{e}}}
 
@@
 \]
-where the bending energy $E_b[\theta(s)] = B/2 \int_0^{s_c} (\theta'(s)-\kappa_o)^2 \ \d s$, the partition
-function $\mathcal{Z} = \int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
+where we define a bending energy for the trajectory captured by the agent's orientation as $E_b[\theta(s)]~=~B/2 \int_0^{s_c} (\theta'(s)-\kappa_o)^2~\d s$, the partition
+function $\mathcal{Z}~=~\int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
 stiffness $B$ captures the constraint for the agent to move forward, $\beta$ is the effective
@@
 target. Instead of solving the partial differential equation for $\P(\varphi, s)$, we can evaluate the
-evolution of the variance of agent's position $\r(s)$ as it marches close to the target.
-
-Assuming that the agent reaching a distance $\sigma$ from the end-point A can reach the
-target, the probability that an agent will reach the target is given by $\P_{\rm OU}(||\r^{n\Delta t} - \r^*|| \leq \sigma)
-= \sigma/\psi(t^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
+evolution of the variance of agent's position $\r(s)$ as it marches close to the target. Let $\ebh_\phi$
+denote the orientation of the agent from $\r(s_c)$ to the target A. We can write the evolution of the variance
+along $\ebh_\perp$ i.e. $\psi(s)=\E[r^2_\perp(s)]$ as $\d{\psi}(s)/\d s = 2 \E [ r_\perp(s) \varphi ]$ for $s>s_c$.
+We can evaluate this evolution to get
+\begin{align}
+    \psi(s) - \psi(s_c) & = \frac{v_o^2 D}{\nu^3} \bigg[ \frac{2 \nu (s-s_c)}{v_o} - \e^{-2 \nu (s-s_c)/v_o} \nonumber \\
+    & \ + 4 \e^{-\nu (s-s_c)/v_o} - 3 \bigg].
+\end{align}
+Assuming that the agent reaching a distance $\sigma$ along $\ebh_\perp$ from the end-point A can reach the
+target (this is associated with the agent's capacity to sense the food source/target),
+the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
+= \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
 \[
-    v_{\rm OU}(\r, \theta) = \int_{0}^{t^*} \int \D[\theta(t)] \P_{\rm OU}[\theta(t)] R(t) \ \d t.
+    v_{\rm OU}(\r) = \int_{s_c}^{s^*} \int \D[\theta(s)] \P_{\rm OU}[\theta(s)] R(s) \ \d s.
 \]
-Let us assume that the reward function for the OU process is $R(t) = \nu[1 - \Theta(||\r(t^*) - \r^*|| - \sigma)]$.
-From this we see that the value of $v_{\rm OU}(\r^c, \theta^c) = \sigma \nu/\psi(t^*)$.
+Let us assume that the reward function for the OU process is $R(s) = \gamma[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$.
+From this we see that the value of $v_{\rm OU}(\r^c) = \sigma \gamma/\psi(s^*)$.
 
@@
 \begin{center}
-\textbf{\large Supplemental Information for ``Branching in soft gels''}\\[.2cm]
-S  Ganga  Prasath$^{1, *}$, Aditi Chakrabarti$^{1, *}$,  and  L  Mahadevan$^{1, 2, 3, 4}$\\[.1cm]
-{\small \itshape ${}^1$School of Engineering and Applied Sciences, Harvard University, Cambridge MA 02138.\\
-${}^2$Department of Physics, Harvard University, Cambridge MA 02138.\\
-${}^3$Center  for  Brain  Science,  Harvard  University,  Cambridge,  MA  02138.\\
-${}^4$Department of Organismic and Evolutionary Biology, Harvard University, Cambridge 02138
+\textbf{\large Supplemental Information for ``Dynamics of innovation in stigmergic collectives''}\\[.2cm]
+S  Ganga  Prasath$^{1}$\\[.1cm]
+{\small \itshape ${}^1$ Department of Applied Mechanics, Indian Institute of Technology Madras, Chennai TN 600 036.
 }

### 2023-05-03 LaTeX/draft.tex
MoreEdits+MathematicaChks

@@
 \maketitle
+Collectives in the wild often have to come up with novel solutions to problems either because
+of uncertainties in the environment that demand new solution or due to high demands in the colony that
+introduces a pressure to innovate. Ants, for egs., use stigmergy to recruit more
+foragers to an identified food source by laying down pheromone along the trajectory they have taken towards the food source.
+Even with pre-treaded paths in place, agents still have to explore to find
+better solutions towards the food source if the demands of the colony are not met.
+In this article, we probe the question of when and how collectives can innovate to find novel solutions
+trajectories for foraging using stigmergy.
+
 \section{Problem setting}
-Let us consider a trail of pheromone $\x^*(s)$ representing the position in 2-D coordinates and parameterized by the
+Social insects such as ants communicate either through pheromone or by antennation. Let us consider a trail of pheromone $\x^*(s)$ representing the position in 2-D coordinates and parameterized by the
 arc-length $s$. The concentration of pheromone is considered continuous along this curve and is given by $c(\x) = c_o \delta(\x-\x^*(s))$.
@@
 direction of motion i.e., its orientation $\ph^t = \{\cos(\theta^t), \sin(\theta^t)\}$, where $[\bullet]^t$ denotes the
-value of the variable at time-instant $t$. The agent can choose the action based on the measurements at each instant of time.
-In this article we are interested in understanding the dynamics of agents that need to innovate in order to find new
-trails over which the collective travels.
+value of the variable at time-instant $t$. The agent chooses its action based on the measurement at each time instant
+which in this case is the concentration of pheromone, $c(\x)$.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{./figs/schematic.pdf}\label{fig:schm1}
+    \caption{Schematic of setup}
+\end{figure*}
 
 \subsection*{Trail following behavior for semi-circular trail}
-Social insects such as ants communicate either through pheromone or by antennation.
-As stigmergy is one of the fundamental mechanisms by which the collective coordinates
-and recruits new workers, we assume that the agent can follow the trail $\x^*(s)$.
-In this article we are interested in understanding how and when an
-agent can innovate from its innate behavior to follow pheromone trails to explore the environment and find novel solutions
-to the navigation problem. In order to understand the dynamics of innovation, we assume that the agent has already
-traversed the trajectory $\x^*(s)$ before.
+
+We know that stigmergy is one of the fundamental mechanisms by which the collective coordinates
+and recruits new workers. With that being the case, the agent can follow the trail $\x^*(s)$ by making
+measurements along its motion trajectory. Since we are interested in understanding how and when an
+agent can innovate from its innate behavior of following pheromone trails to explore the environment and find novel solutions
+to the navigation problem, we assume that the agent has already traversed the trajectory $\x^*(s)$ before.
 This intrinsic behavior is denoted by the policy $\pi_{\rm int}(\r)$ which provides the
 appropriate angle $\theta^t$ required to follow the trail when the agent is at location
-$\r^t$ after the measurment that the agent makes. In order to consider a specific trail
-of the initial pheromone trail, let us assume that the trail is a simple semi-circle with
-functional form $\xst(s) = \{ a \cos(s/a), a \sin(s/a) \}$. For this particular case,
-the policy can be explicitly written as
+$\r^t$ after the measurment that the agent makes. We consider the trail to be a simple semi-circle
+with functional form $\xst(s) = \{ a \cos(s/a), a \sin(s/a) \}$ (see Fig.~\ref{fig:schm1}$(b)$). For this particular case,
+the intrinsic trail following policy can be explicitly written as
 \begin{align}
@@
 \end{align}
-where $\zeta^t$ is the fluctuation/error in the measurements that the agent makes, $l$ is the step-length
+where $\zeta^t$ is the noise/error in the agent's measurement, $l$ is the step-length
 of the agent. In the limit when the step-length is small, we can write the continuous version
 of the motion as $\dot{\r} = v_o \ph$. Without loss of generality, we can write this as
-$\d\r/\d s = \ph$ where $s = v_o t$ and we assume that the distribution of the orientation of the agent is distributed as though
-it is a worm-like chain polymer whose orientation distribution is given by
+$\d\r/\d s = \ph$ where $s = v_o t$. Moreover when the orientation of the agent
+along its trajectory is distributed as though it were a worm-like chain polymer,
+the functional form of orientation distribution is
 \[
-    \P[\theta(s)] = \frac{1}{\mathcal{Z}}\exp{(-\beta E_b)}
+    \P[\theta(s)] = \frac{1}{\mathcal{Z}}\exp{(-\beta E_b)},
 \]
@@
 function $\mathcal{Z}~=~\int \mathcal{D}[\theta(s)] \exp\{ {-\beta E_b[\theta(s)]} \}$ and the bending
-stiffness $B$ captures the constraint for the agent to move forward, $\beta$ is the effective
+stiffness $B$ captures the constraint for the agent to move forward towards the target A (ref Fig.~\ref{fig:schm1}$(b)$), $\beta$ is the effective
 temperature capturing the fluctuations in the trajectory of agent. We also have
-boundary conditions $\theta(0) = \pi/2$ and $\theta'(0) = \kappa_o = a^{-1}$. Here $s_c$ is the coordinate
-along the semi-circle at which the agent bifurcates towards the target location A using a new trajectory.
+boundary conditions $\theta(0) = \pi/2$ (due to initial orientation of the agent) and $\theta'(0) = \kappa_o = a^{-1}$ (due to curvature of the trail).
+Here $s_c$ is the coordinate along the semi-circle at which the agent bifurcates towards the target location A using a new trajectory.
 Our motivation is to identify this location using the Bellman equation which will be discussed later.
 Now in order to evaluate the quality of the intrinsic policy we introduce the notion of value function
-of this policy $\pi_{\rm int}(\r)$,denoted by $v_{\rm int}(\r^t)$. The value function provides the cummulative
-future reward that the agent will acquire when following the policy $\pi_{\rm int}(\r)$. This can be written as
+of this policy $\pi_{\rm int}(\r)$, denoted by $v_{\rm int}(\r)$. The value function provides the cummulative
+future reward that the agent will acquire from location $\r$ at time $t$
+when following the policy $\pi_{\rm int}(\r)$. This can be written as
 \[
@@
 \]
-We can write this in path-integral form as
+We can write this in path-integral form as (with $\gamma = 1$)
 \begin{align}
@@
 \[
-    v_{\rm int}(\r) = \frac{\phi^*}{\kappa_o} [ \e^{-\kappa_os/\phi^*} - \e^{-\kappa_o s_c/\phi^*}].
+    v_{\rm int}(\r) = \frac{\phi^*}{\kappa_o} [ \e^{-\kappa_o s/\phi^*} - \e^{-\kappa_o s_c/\phi^*}].
 \]
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{./figs/schematic.pdf}\label{fig:schm1}
-    \caption{Schematic of setup}
-\end{figure*}
 
@@
 Though the agent can take a variety of exploration strategies, we assume that the policy resembles
-an Ornstein-Uhlenbeck (OU) process. This explorative policy $\pi_{\OU}(\r)$ can be written through $\P(\theta, t)$
-the probability that the agent will take a particular angle $\theta$ at time $t$ or distance $s$. However since the orientation of the
-agent is dependent on $\r(s_c)$, the angle between agent location $\r(s_c)$ to target A can be denoted
-as $\phi$ and $\theta=\phi+\varphi$ where $\varphi$ is the fluctuation around this orientation that follows
-an OU-process. The probability $\P(\varphi, s)$ can then be written as
+an Ornstein-Uhlenbeck (OU) process in the agent's orientation. This fits well with the behavior of an insect such as an ant because
+ants often explore using their orientaiton after making measurements with their antennae.
+This explorative policy $\pi_{\OU}(\r)$ can be written through $\P(\theta, t)$
+the probability that the agent will take a particular angle $\theta$ at time $t$ or equivalently at distance $s$.
+However, since the orientation of the agent is dependent on $\r(s_c)$, the angle between agent location $\r(s_c)$
+to target A can be denoted as $\phi$ and $\theta=\phi+\varphi$ where $\varphi$ is the fluctuation around this
+orientation that follows an OU-process. The probability $\P(\varphi, s)$ can then be written as
 \[
@@
 and $D$ represents the fluctuation in orientation, which we will soon assume to be large when $\r(s_c)$
-is farther from the target A. We assume that the agent reaching a distance $\sigma$ from the end-point A can reach the
-target and thus the reward is a function of the propagation of variance in agent's position towards the
+is farther from the target A. Often in navigating insects, agents can quickly detect food source when they
+are close to the food source. We capture this by assuming that after reaching a distance $\sigma$ from the end-point A,
+agents can reach the target and thus the reward is a function only of the variance in agent's position towards the
 target. Instead of solving the partial differential equation for $\P(\varphi, s)$, we can evaluate the
@@
 \end{align}
-Assuming that the agent reaching a distance $\sigma$ along $\ebh_\perp$ from the end-point A can reach the
-target (this is associated with the agent's capacity to sense the food source/target),
-the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
-= \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ using the formula
+As already mentioned, the agent reaching a distance $\sigma$ along $\ebh_\perp$ from the end-point A can reach the
+target, the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
+= \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
 \[
@@
 \]
-Let us assume that the reward function for the OU process is $R(s) = \gamma[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$.
-From this we see that the value of $v_{\rm OU}(\r^c) = \sigma \gamma/\psi(s^*)$.
+The reward function for the OU process takes a simple form $R(s) = \mu[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$,
+which simply means that the agent reaching with a distance $\sigma$ gets a reward $\mu$.
+From this we see that the value of $v_{\rm OU}(\r^c) = \sigma \mu/\psi(s^*)$.
+
+\section*{Value function of combined policy}
+The total value function for the agent's policy is $v(\r, \theta) = v_{\rm int}(\r, \theta) + v_{\rm OU}(\r^c, \theta^c)$.
+The agent is trying to maximize this value function from a given state, $\r^t, \theta^t$. The final expression
+for the value function is
+\[
+    v(\r; s_c, D) = \frac{\sigma \mu}{\psi(s^*; s_c, D)} + \frac{\phi^*}{\kappa_o} [ \e^{-\kappa_os/\phi^*} - \e^{-\kappa_o s_c/\phi^*}],
+\]
+the region around the target location A. In this equation, the diffusion coefficient
+representing the fluctuations in the dynamics of the agent capturing the error in memory
+of target location A. We assume that the fluctuation $D$ is larger farther from the target location.
 

### 2023-05-04 LaTeX/draft.tex
ValueFunctionComplete

@@
 \def\ebh{\hat{\mathbf{e}}}
+\def\Pe{\mathsf{Pe}}
 
@@
 \end{align}
-As already mentioned, the agent reaching a distance $\sigma$ along $\ebh_\perp$ from the end-point A can reach the
-target, the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
+Before we move further with the calculation, it is cogent to work in non-dimensional numbers. In the OU-process
+we have two length-scales: $l \sim (v_o/\nu), (v_o/D)$, one due to \textit{inertia} and other due to \textit{diffusion}.
+The ratio of these two is nothing but the Peclet number, $\Pe = (D/\nu)$. We choose $l_\nu = (v_o/\nu)$ as the length-scale
+for further non-dimensionalization such that $s \rightarrow l_\nu \hat{s}$. We drop the hat in what follows and we get
+the non-dimensional variance, $\psi(s)$ (after setting $\psi(s_c) =0$) as a function of non-dimensional distance, $s$ as
+\begin{align}
+    \psi(s)& = \Pe \bigg[ 2 (s-s_c) - \e^{-2 (s-s_c)} + 4 \e^{-(s-s_c)} - 3 \bigg].
+\end{align}
+As already mentioned, the agent reaching a distance $\sigma$ (appropriately non-dimensionalized with $l_\nu$) along $\ebh_\perp$ from the end-point A
+can reach the target, the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
 = \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
@@
 \[
-    v(\r; s_c, D) = \frac{\sigma \mu}{\psi(s^*; s_c, D)} + \frac{\phi^*}{\kappa_o} [ \e^{-\kappa_os/\phi^*} - \e^{-\kappa_o s_c/\phi^*}],
+    v(\r; s_c, D) = \frac{\sigma \mu}{\psi(s^*; s_c, \Pe)} + \frac{\phi^*}{\kappa} [ \e^{-\kappa s/\phi^*} - \e^{-\kappa s_c/\phi^*}],
 \]
-the region around the target location A. In this equation, the diffusion coefficient
-representing the fluctuations in the dynamics of the agent capturing the error in memory
-of target location A. We assume that the fluctuation $D$ is larger farther from the target location.
+where we have introduced the non-dimensional number, $\kappa = \kappa_o l_\nu = (\kappa_o v_o/\nu)$.
+In this equation, the Peclet number $\Pe$ captures the fluctuations in the dynamics of the agent
+associated with the error in the homing vector. We assume that this diffusion is larger farther
+from the target location and takes the functional form $\Pe \sim 1/s$. Moreover the reward that the agent recieves
+when the bifurcation point is close to the target location needs to be smaller and thus $\mu = (\pi\kappa^{-1} -s_c)$.
+One last detail that needs to be specified to make the value function complete is to specify $s^*$. It is straight
+forward to show that from $s_c$, the target A is a distance $2 \cos(s_c\kappa/2)\kappa^{-1}$ which makes
+$s^* = s_c + 2 \cos(s_c\kappa/2)\kappa^{-1}$.
+
+\subsection*{Single agent learning}
 

### 2023-05-07 LaTeX/draft.tex
HomeEdits

@@
 The ratio of these two is nothing but the Peclet number, $\Pe = (D/\nu)$. We choose $l_\nu = (v_o/\nu)$ as the length-scale
-for further non-dimensionalization such that $s \rightarrow l_\nu \hat{s}$. We drop the hat in what follows and we get
-the non-dimensional variance, $\psi(s)$ (after setting $\psi(s_c) =0$) as a function of non-dimensional distance, $s$ as
+for further non-dimensionalization such that $s \rightarrow l_\nu \hat{s}, t \rightarrow \nu^{-1} \hat{t}$. We drop the hat in what follows and we get
+the non-dimensional variance, $\psi(s)$ (after setting $\psi(s_c) =0$) as a function of non-dimensional distance $s$ to be
 \begin{align}
@@
 As already mentioned, the agent reaching a distance $\sigma$ (appropriately non-dimensionalized with $l_\nu$) along $\ebh_\perp$ from the end-point A
-can reach the target, the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
+can reach the target, then the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
 = \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
@@
 The reward function for the OU process takes a simple form $R(s) = \mu[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$,
-which simply means that the agent reaching with a distance $\sigma$ gets a reward $\mu$.
-From this we see that the value of $v_{\rm OU}(\r^c) = \sigma \mu/\psi(s^*)$.
+which simply means that the agent reaching within a distance $\sigma$ gets a reward $\mu$.
+From this we see that the value function from $\r$ to the end of the trajectory, $v_{\rm OU}(\r^c) = \sigma \mu/\psi(s^*)$.
 
 \section*{Value function of combined policy}
-The total value function for the agent's policy is $v(\r, \theta) = v_{\rm int}(\r, \theta) + v_{\rm OU}(\r^c, \theta^c)$.
+The total value function for the agent's policy is simply $v(\r, \theta) = v_{\rm int}(\r, \theta) + v_{\rm OU}(\r^c, \theta^c)$.
 The agent is trying to maximize this value function from a given state, $\r^t, \theta^t$. The final expression

### 2023-05-08 LaTeX/draft.tex
RewardDone

@@
 \end{align}
+The non-dimensional Langevin dynamics that results in this evolution of the variance is $\d \varphi/\d s = - \varphi + \sqrt{2 \Pe} \eta(s)$.
 As already mentioned, the agent reaching a distance $\sigma$ (appropriately non-dimensionalized with $l_\nu$) along $\ebh_\perp$ from the end-point A

### 2023-05-10 LaTeX/draft.tex
EpsilonGreedyWorking

@@
 \subsection*{Single agent learning}
+Now that we have the environment and the agent behavior laid out, we look at the learning dynamics of the agent
+when it has no feedback through pheromone but still tries to perform the distance-accuracy tradeoff. The agent can
+make this tradeoff by exploring the environment as well as exploiting its intrinsic ability to follow pheromone to
+accumulate reward. This is established through an $\epsilon$-greedy policy and we look at the evolution of the value
+function as the agent so learns by following this strategy.
+
+Mention that the TD(0)-learning results in
+\[
+    v(s_c) \leftarrow (1-\alpha) v(s_c) + \alpha r(s^*)
+\]
+as $v(s^*) = 0$.
 

### 2023-05-10 plotFns.py

@@ -1,0 +1,480 @@
+# Plot functions
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib.patheffects as path_effects
+
+def plotDef(rmax, p, xphm, yphm, phm):
+    global colors
+    mpl.rcParams.update(mpl.rcParamsDefault)
+    plt.rc("font", family="Helvetica Neue")
+    plt.rc("xtick", labelsize="medium")
+    linewidth = 1.2
+    linestyle = "-"
+    label_size = 15
+    color = '#EE3984'
+    lcolor = '#FFC708'
+
+    fig, ax = plt.subplots(1,3,figsize=(12,4), gridspec_kw={'width_ratios': [1, 1, 1]}, dpi=150)
+    ax[0].contourf(xphm, yphm, phm.c, cmap='Purples')
+    thetPlt = np.linspace(0., np.pi, 200)
+    ax[0].plot(p.a*np.cos(thetPlt), p.a*np.sin(thetPlt), c='white', zorder=1)
+
+    ax[0].set_xlabel(r'$x$',usetex=True, fontsize=label_size)
+    ax[0].set_ylabel(r'$y$',usetex=True,rotation=0, fontsize=label_size)
+    ax[0].xaxis.set_label_coords(0.5,-0.1)
+    ax[0].yaxis.set_label_coords(-0.15,0.4)
+    ax[0].set_aspect(1)
+    ax[0].set_ylim([-0.5,1.5])
+    ax[0].set_xlim([-1.5,1.5])
+
+
+    ax[1].set_xlabel(r'$t$',usetex=True, fontsize=label_size)
+    ax[1].set_ylabel(r'$R(t)$',usetex=True,rotation=0, fontsize=label_size)
+    ax[1].set_xlim([0., 10.])
+    ax[1].set_ylim([0.,1.0])
+
+    ax[1].xaxis.set_label_coords(0.5,-0.15)
+    ax[1].yaxis.set_label_coords(-0.1,0.4)
+
+    ax[1].spines["right"].set_visible(False)
+    ax[1].spines["top"].set_visible(False)
+    ax[1].spines["left"].set_position(("data", 0))
+    ax[1].spines["bottom"].set_position(("data", 0))
+    ax[1].set_aspect(7)
+
+    ax[2] = plt.subplot(133, projection='polar')
+    ax[2].set_thetalim(0, np.pi)
[…]

### 2023-05-10 prmts.py

@@ -1,0 +1,42 @@
+import numpy as np
+import scipy as sp
+# Parameters in the model
+class params():
+    def __init__(self):
+        super().__init__()
+        self.nSteps = 800 #number of steps of each run
+        self.nEpchs = 2000 #number of epochs
+        self.a = 1.0 #radius of semi-circle
+        self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
+        
+        # Pheromone parameters
+        self.nMsh = 500 #number of points in 
+        self.kM = 1 #rate of pheromone decay
+        self.kP = 0.1 #rate of pheromone generation
+        
+        # Agent dynamics parameters
+        self.pt = 0.05 #initial pheromone trail half-thickness
+        self.sz = 0.02 #agent size (radius)
+        self.dt = 1e-2 #time-step size
+        self.l = 5e-3 #length travelled in 1 time-step
+        self.vo = self.l/self.dt #effective speed of motion
+        self.nu = 50 #orientation relaxation rate (s^-1)
+        # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
+        # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
+        self.diffCst = 50 #noise diffusion coefficient (s^-1)
+        self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
+        self.pe = self.diff/self.nu #peclet number
+        
+        # Reward/Learning parameters
+        self.alpha = 0.9 #learning rate
+        self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
+        self.sigma = 0.1 #radius of region near target within which agent succeeds
+        self.nPtn = 20 #number of division of \phi over which value function is define
+        self.epsilon = 0.15 #constant for epsilon-greedy strategy
+        
+        # Initial conditions
+        self.rInit = np.array([self.a + np.random.uniform(-self.pt, self.pt), 0.0]) #initial location of agent
+        self.tgt = np.array([-self.a, 0.0]) #target location of agent
+        self.thetInit = 0.5*np.pi #initial orientation of agent
+        # self.rng = random.PRNGKey(1) #initial random number
+        self.tgtPhi = 0.0 #target orientation from current location

### 2023-05-10 prmts.py
MultiAgentsStrt

@@ -5,7 +5,7 @@ class params():
     def __init__(self):
         super().__init__()
         self.nSteps = 800 #number of steps of each run
-        self.nEpchs = 2000 #number of epochs
+        self.nEpchs = 10000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
         

### 2023-05-11 prmts.py
TDLearningFile

@@ -5,7 +5,7 @@ class params():
     def __init__(self):
         super().__init__()
         self.nSteps = 800 #number of steps of each run
-        self.nEpchs = 10000 #number of epochs
+        self.nEpchs = 5000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
         
@@ -31,8 +31,8 @@ class params():
         self.alpha = 0.9 #learning rate
         self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
         self.sigma = 0.1 #radius of region near target within which agent succeeds
-        self.nPtn = 20 #number of division of \phi over which value function is define
-        self.epsilon = 0.15 #constant for epsilon-greedy strategy
+        self.nPtn = 40 #number of division of \phi over which value function is define
+        self.epsilon = 0.8 #constant for epsilon-greedy strategy
         
         # Initial conditions
         self.rInit = np.array([self.a + np.random.uniform(-self.pt, self.pt), 0.0]) #initial location of agent

### 2023-05-11 plotFns.py
ValueFnConvergenceUpdate

@@ -32,7 +32,7 @@ def plotDef(rmax, p, xphm, yphm, phm):
     ax[1].set_xlabel(r'$t$',usetex=True, fontsize=label_size)
     ax[1].set_ylabel(r'$R(t)$',usetex=True,rotation=0, fontsize=label_size)
     ax[1].set_xlim([0., 10.])
-    ax[1].set_ylim([0.,1.0])
+    ax[1].set_ylim([0.,5.0])
 
     ax[1].xaxis.set_label_coords(0.5,-0.15)
     ax[1].yaxis.set_label_coords(-0.1,0.4)
@@ -41,7 +41,7 @@ def plotDef(rmax, p, xphm, yphm, phm):
     ax[1].spines["top"].set_visible(False)
     ax[1].spines["left"].set_position(("data", 0))
     ax[1].spines["bottom"].set_position(("data", 0))
-    ax[1].set_aspect(7)
+    ax[1].set_aspect(2)
 
     ax[2] = plt.subplot(133, projection='polar')
     ax[2].set_thetalim(0, np.pi)
@@ -143,22 +143,23 @@ def plotDef(rmax, p, xphm, yphm, phm):
     return fig, ax
 
 
-def plotDat(fig, ax, p, rxFull, ryFull, phxFull, phyFull, rwdFull, valFn):
+def plotDat(fig, ax, p, rxFull, ryFull, phxFull, phyFull, rwdFull, valFn, alph):
     # Plot data
     color = '#EE3984'
     lcolor = '#FFC708'
     global colors
     ax[0].scatter(rxFull, ryFull,
-            s=25,
+            s=15,
             color=color,
             # edgecolors='k',
             linewidth=0.5,
-            alpha=0.005,
+            alpha=alph/500,
             zorder=2)
     
     t = np.arange(np.size(rwdFull))*p.dt
     ax[1].plot(t, rwdFull,
-               alpha=0.1,
+            #    alpha=0.1,
+               alpha=alph/10,
                color=lcolor)
 
     n = 2*p.nPtn + 1
@@ -177,9 +178,9 @@ def plotDat(fig, ax, p, rxFull, ryFull, phxFull, phyFull, rwdFull, valFn):
     # R[-1] = R[0]
     # R[-2] = R[0]
     
-    ax[2].fill(phiPlt, R, color="C1", zorder=150, alpha=0.005)
[…]

### 2023-05-11 prmts.py

@@ -5,7 +5,7 @@ class params():
     def __init__(self):
         super().__init__()
         self.nSteps = 800 #number of steps of each run
-        self.nEpchs = 5000 #number of epochs
+        self.nEpchs = 10000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
         
@@ -28,7 +28,7 @@ class params():
         self.pe = self.diff/self.nu #peclet number
         
         # Reward/Learning parameters
-        self.alpha = 0.9 #learning rate
+        self.alpha = 0.7 #learning rate
         self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
         self.sigma = 0.1 #radius of region near target within which agent succeeds
         self.nPtn = 40 #number of division of \phi over which value function is define

### 2023-05-11 plotFns.py
MultiAgentBasic

@@ -304,7 +304,130 @@ def plotRadial(rmax, p, valFn):
     ax.plot(phiPlt, R, color="C1", zorder=250, linewidth=1, alpha=1.)
     
     return fig, ax
-    
+
+def plotDefMul(p, xphm, yphm, phm):
+    global colors
+    mpl.rcParams.update(mpl.rcParamsDefault)
+    plt.rc("font", family="Helvetica Neue")
+    plt.rc("xtick", labelsize="medium")
+    linewidth = 1.2
+    linestyle = "-"
+    label_size = 15
+    color = '#EE3984'
+    lcolor = '#FFC708'
+
+    fig, ax = plt.subplots(1,2,figsize=(8,4), gridspec_kw={'width_ratios': [1, 1]}, dpi=150)
+    ax[0].contourf(xphm, yphm, phm.c, cmap='Purples')
+    thetPlt = np.linspace(0., np.pi, 200)
+    ax[0].plot(p.a*np.cos(thetPlt), p.a*np.sin(thetPlt), c='white', zorder=1)
+
+    ax[0].set_xlabel(r'$x$',usetex=True, fontsize=label_size)
+    ax[0].set_ylabel(r'$y$',usetex=True,rotation=0, fontsize=label_size)
+    ax[0].xaxis.set_label_coords(0.5,-0.1)
+    ax[0].yaxis.set_label_coords(-0.15,0.4)
+    ax[0].set_aspect(1)
+    ax[0].set_ylim([-0.5,1.5])
+    ax[0].set_xlim([-1.5,1.5])
+
+
+    ax[1].set_xlabel(r'$t$',usetex=True, fontsize=label_size)
+    ax[1].set_ylabel(r'$R(t)$',usetex=True,rotation=0, fontsize=label_size)
+    ax[1].set_xlim([0., 10.])
+    ax[1].set_ylim([0.,5.0])
+
+    ax[1].xaxis.set_label_coords(0.5,-0.15)
+    ax[1].yaxis.set_label_coords(-0.1,0.4)
+
+    ax[1].spines["right"].set_visible(False)
+    ax[1].spines["top"].set_visible(False)
+    ax[1].spines["left"].set_position(("data", 0))
+    ax[1].spines["bottom"].set_position(("data", 0))
+    ax[1].set_aspect(1)
+
+    return fig, ax
+
+def plotMultAgt(fig, ax, p, rxFull, ryFull, phxFull, phyFull, alph):
+    # Plot data
+    color = '#EE3984'
[…]

### 2023-05-11 prmts.py

@@ -15,7 +15,7 @@ class params():
         self.kP = 0.1 #rate of pheromone generation
         
         # Agent dynamics parameters
-        self.pt = 0.05 #initial pheromone trail half-thickness
+        self.pt = 0.04 #initial pheromone trail half-thickness
         self.sz = 0.02 #agent size (radius)
         self.dt = 1e-2 #time-step size
         self.l = 5e-3 #length travelled in 1 time-step

### 2023-06-01 plotFns.py
SmallChanges

@@ -369,7 +369,8 @@ def plotMultAgt(fig, ax, p, rxFull, ryFull, phxFull, phyFull, alph):
            phxFull, phyFull,
            scale_units='xy',
            color=acolor,
-           scale=10.2)
+        #    scale=10.2)
+           scale=20.2)
     return fig, ax
 
 def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
@@ -393,7 +394,7 @@ def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
                 thetPlt = np.linspace(0., np.pi, 200)
                 ax[0].plot(p.a*np.cos(thetPlt), p.a*np.sin(thetPlt), c='white', zorder=1)
                 ax[0].scatter(rxFull[ind], ryFull[ind],
-                        s=30,
+                        s=10,
                         color=color,
                         edgecolors='k',
                         linewidth=0.5,
@@ -403,7 +404,7 @@ def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
                         phxFull[ind], phyFull[ind],
                         scale_units='xy',
                         color=acolor,
-                        scale=10.2)
+                        scale=20.2)
                 ax[0].set_xlabel(r'$x$',usetex=True, fontsize=label_size)
                 ax[0].set_ylabel(r'$y$',usetex=True,rotation=0, fontsize=label_size)
                 ax[0].xaxis.set_label_coords(0.5,-0.05)

### 2023-06-02 multPrmts.py
MultiAgtUpdates

@@ -1,0 +1,44 @@
+import numpy as np
+import scipy as sp
+# Parameters in the model
+class params():
+    def __init__(self) -> None:
+        super().__init__()
+        self.nSteps = 900 #number of steps of each run
+        self.nAgnts = 10 #number of agents
+        self.nForag = 1 #number of foragers
+        # self.nEpchs = 250 #number of epochs
+        self.a = 1.0 #radius of semi-circle
+        self.nIntSteps = int(0.5*self.nSteps) #number of steps of intrinsic policy
+        self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
+        
+        # Pheromone parameters
+        self.nMsh = 500 #number of points in 
+        self.kM = 1 #rate of pheromone decay
+        self.kP = 0.1 #rate of pheromone generation
+        
+        # Agent dynamics parameters
+        self.pt = 0.08 #initial pheromone trail thickness
+        self.sz = 0.01 #agent size (radius)
+        self.dt = 1e-2 #time-step size
+        self.l = 0.5e-2 #length travelled in 1 time-step
+        self.vo = self.l/self.dt #effective speed of motion
+        self.nu = 50 #orientation relaxation rate (s^-1)
+        self.diff = 50e-1 #noise diffusion coefficient (s^-1)
+        self.diffCst = 50 #noise diffusion coefficient (s^-1)
+        # self.pe = self.diff/self.nu #peclet number
+        
+        # Reward parameters
+        self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
+        self.sigma = 0.1 #radius of region near target within which agent succeeds
+        self.nPtn = 40 #number of division of \phi over which value function is define
+        
+        # Initial conditions
+        phiInit = np.random.uniform(0, np.pi, self.nAgnts)
+        self.rInit = self.a*np.array([np.cos(phiInit), np.sin(phiInit)])
+        self.tgt = np.array([-self.a, 0.0]) #target location of agent
+        self.strt = np.array([self.a, 0.0]) #initial location of agent
+        self.thetInit = np.random.uniform(0, 2*np.pi, self.nAgnts)
+
+        self.rFgInit = self.strt
+        self.thetFgInit = np.pi*0.5

### 2023-06-02 plotFns.py

@@ -359,7 +359,8 @@ def plotMultAgt(fig, ax, p, rxFull, ryFull, phxFull, phyFull, alph):
     lcolor = '#FFC708'
     acolor = '#7A7A7A'
     ax[0].scatter(rxFull, ryFull,
-            s=10,
+            # s=10,
+            s=4,
             color=color,
             edgecolors='k',
             linewidth=0.5,
@@ -370,7 +371,28 @@ def plotMultAgt(fig, ax, p, rxFull, ryFull, phxFull, phyFull, alph):
            scale_units='xy',
            color=acolor,
         #    scale=10.2)
-           scale=20.2)
+           scale=30.2)
+    return fig, ax
+
+def plotMultFg(fig, ax, p, rxFull, ryFull, phxFull, phyFull, alph):
+    # Plot data
+    color = '#38BDF2'
+    lcolor = '#FFC708'
+    acolor = '#7A7A7A'
+    ax[0].scatter(rxFull, ryFull,
+            # s=10,
+            s=4,
+            color=color,
+            edgecolors='k',
+            linewidth=0.5,
+            alpha=alph,
+            zorder=2)
+    ax[0].quiver(rxFull, ryFull,
+           phxFull, phyFull,
+           scale_units='xy',
+           color=acolor,
+        #    scale=10.2)
+           scale=30.2)
     return fig, ax
 
 def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
@@ -394,7 +416,8 @@ def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
                 thetPlt = np.linspace(0., np.pi, 200)
                 ax[0].plot(p.a*np.cos(thetPlt), p.a*np.sin(thetPlt), c='white', zorder=1)
                 ax[0].scatter(rxFull[ind], ryFull[ind],
-                        s=10,
+                        # s=10,
+                        s=4,
                         color=color,
                         edgecolors='k',
                         linewidth=0.5,
@@ -404,7 +427,8 @@ def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
                         phxFull[ind], phyFull[ind],
                         scale_units='xy',
                         color=acolor,
-                        scale=20.2)
[…]

### 2023-06-02 plotFns.py
MinorEdits

@@ -460,3 +460,84 @@ def plotVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull):
                 plt.close()
                 fInd += 1
         return
+
+def plotMulVidFn(p, xphm, yphm, phm, rxFull, ryFull, phxFull, phyFull, rfgxFull, rfgyFull, phfgxFull, phfgyFull):
+        mpl.rcParams.update(mpl.rcParamsDefault)
+        plt.rc("font", family="Helvetica Neue")
+        plt.rc("xtick", labelsize="medium")
+        linewidth = 1.2
+        linestyle = "-"
+        label_size = 15
+        color = '#EE3984'
+        fgColor = '#38BDF2'
+        lcolor = '#FFC708'
+        acolor = '#7A7A7A'
+
+        t = np.arange(np.size(rxFull))*p.dt
+        rwdFull = np.zeros(np.size(rxFull))
+        fInd = 0
+        for ind in np.arange(0,rxFull.shape[0],2):
+        # for ind in range(10):
+                fig, ax = plt.subplots(1,2,figsize=(9,9), gridspec_kw={'width_ratios': [1, 1]}, dpi=150)
+                ax[0].contourf(xphm, yphm, phm.c, cmap='Purples')
+                thetPlt = np.linspace(0., np.pi, 200)
+                ax[0].plot(p.a*np.cos(thetPlt), p.a*np.sin(thetPlt), c='white', zorder=1)
+                ax[0].scatter(rxFull[ind], ryFull[ind],
+                        # s=10,
+                        s=4,
+                        color=color,
+                        edgecolors='k',
+                        linewidth=0.5,
+                        alpha=1,
+                        zorder=2)
+                ax[0].quiver(rxFull[ind], ryFull[ind],
+                        phxFull[ind], phyFull[ind],
+                        scale_units='xy',
+                        color=acolor,
+                        # scale=20.2)
+                         scale=30.2)
+                
+                ax[0].scatter(rfgxFull[ind], rfgyFull[ind],
+                        s=4,
+                        color=fgColor,
+                        edgecolors='k',
+                        linewidth=0.5,
+                        alpha=1,
+                        zorder=2)
+                ax[0].quiver(rfgxFull[ind], rfgyFull[ind],
+                        phfgxFull[ind], phfgyFull[ind],
+                        scale_units='xy',
+                        color=acolor,
[…]

### 2023-06-13 LaTeX/draft.tex
Remark
MoreEdits

@@
 \]
-as $v(s^*) = 0$.
+as $v(s^*) = 0$. Mention that we don't need $q(s, a)$ as the value function is parameterized in a fashion to account for
+actions.
 
@@
 while the co-variance dynamics is given by $\dot{\zeta}(t) = -2\alpha \zeta(t) + 2D$ with
-$\zeta = \E [  {(\theta(t) - h(t))}^2 ] = \E [  \theta^2(t) ] - h{(t)}^2$.
+$\zeta = \E [  {(\theta(t) - h(t))}^2 ] = \E [  \theta^2(t) ] - h{(t)}^2$. We can solve this to get
+the time-evolution of the co-variance as $\zeta(t) = D/\alpha(1-\e^{-2\alpha t})$.
 
@@
 \end{align}
-This second order equation needs to boundary condition. We require $y(0) = 0$ and $\yd (0) = 0$ which
+This second order equation needs two boundary condition. We require $y(0) = 0$ and $\yd (0) = 0$ which
 comes from setting the initial orientation to be 0. We are interested in the evolution of the mean and

### 2023-06-15 LaTeX/draft.tex
CodeUpdates

@@
 \]
-We can write this in path-integral form as (with $\gamma = 1$)
+We can write this in path-integral form as (with $\gamma \approx 1$)
 \begin{align}
@@
 The non-dimensional Langevin dynamics that results in this evolution of the variance is $\d \varphi/\d s = - \varphi + \sqrt{2 \Pe} \eta(s)$.
-As already mentioned, the agent reaching a distance $\sigma$ (appropriately non-dimensionalized with $l_\nu$) along $\ebh_\perp$ from the end-point A
-can reach the target, then the probability that an agent will reach the target is given by $\P(||r_\perp^{s^*} - r^*|| \leq \sigma)
-= \sigma/\psi(s^*)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
+As already mentioned, the agent reaching a distance $\sigma$ (appropriately non-dimensionalized with $l_\nu$) from the end-point A
+can reach the target, then the probability that an agent will reach the target is given by $\P(||\r(s) - \r^*|| \leq \sigma)
+= 2\int_0^\sigma \mu \sqrt{\sigma^2-\zeta^2} \ \d \zeta/\psi(s^*-\zeta)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
 \[
@@
 \]
-The reward function for the OU process takes a simple form $R(s) = \mu[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$,
+% The reward function for the OU process takes a simple form $R(s) = \mu[1 - \Theta(||r_\perp(s^*) - r^*|| - \sigma)]$,
+The reward function for the OU process takes a simple form $R(s) = \mu[1 - \Theta(||\r(s^*) - \r^*||^2 - \sigma^2)]$,
 which simply means that the agent reaching within a distance $\sigma$ gets a reward $\mu$.
 From this we see that the value function from $\r$ to the end of the trajectory, $v_{\rm OU}(\r^c) = \sigma \mu/\psi(s^*)$.
+Mention the limits of $\sigma$ set by the criteria that $\psi(s) > 0$.
 

### 2023-06-15 prmts.py

@@ -4,13 +4,13 @@ import scipy as sp
 class params():
     def __init__(self):
         super().__init__()
-        self.nSteps = 800 #number of steps of each run
+        self.nSteps = 1000 #number of steps of each run
         self.nEpchs = 10000 #number of epochs
         self.a = 1.0 #radius of semi-circle
-        self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
+        self.nSc = int(0.3*self.nSteps) #number of steps of intrinsic policy
         
         # Pheromone parameters
-        self.nMsh = 500 #number of points in
+        self.nMsh = 500 #number of points in grid
         self.kM = 1 #rate of pheromone decay
         self.kP = 0.1 #rate of pheromone generation
         

### 2023-06-16 prmts.py
MinorEdits

@@ -5,9 +5,9 @@ class params():
     def __init__(self):
         super().__init__()
         self.nSteps = 1000 #number of steps of each run
-        self.nEpchs = 10000 #number of epochs
+        self.nEpchs = 5000 #number of epochs
         self.a = 1.0 #radius of semi-circle
-        self.nSc = int(0.3*self.nSteps) #number of steps of intrinsic policy
+        self.nSc = int(0.05*self.nSteps) #number of steps of intrinsic policy
         
         # Pheromone parameters
         self.nMsh = 500 #number of points in grid
@@ -23,7 +23,7 @@ class params():
         self.nu = 50 #orientation relaxation rate (s^-1)
         # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
-        self.diffCst = 50 #noise diffusion coefficient (s^-1)
+        self.diffCst = 5. #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
         

### 2023-06-16 plotFns.py
Chngs

@@ -164,8 +164,8 @@ def plotDat(fig, ax, p, rxFull, ryFull, phxFull, phyFull, rwdFull, valFn, alph):
     ax[1].plot(t, rwdFull,
             #    alpha=0.1,
                alpha=alph/10,
-               color=colors[rcInd])
-            #    color=lcolor)
+            #    color=colors[rcInd])
+               color=lcolor)
 
     n = 2*p.nPtn + 1
     phiPlt = np.linspace(0, np.pi, n)

### 2023-06-16 prmts.py

@@ -5,7 +5,7 @@ class params():
     def __init__(self):
         super().__init__()
         self.nSteps = 1000 #number of steps of each run
-        self.nEpchs = 5000 #number of epochs
+        self.nEpchs = 8000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.05*self.nSteps) #number of steps of intrinsic policy
         
@@ -20,10 +20,10 @@ class params():
         self.dt = 1e-2 #time-step size
         self.l = 5e-3 #length travelled in 1 time-step
         self.vo = self.l/self.dt #effective speed of motion
-        self.nu = 50 #orientation relaxation rate (s^-1)
+        self.nu = 10 #orientation relaxation rate (s^-1)
         # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
-        self.diffCst = 5. #noise diffusion coefficient (s^-1)
+        self.diffCst = 0.1 #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
         

### 2023-06-16 prmts.py
Chng

@@ -4,7 +4,8 @@ import scipy as sp
 class params():
     def __init__(self):
         super().__init__()
-        self.nSteps = 1000 #number of steps of each run
+        # self.nSteps = 1000 #number of steps of each run
+        self.nSteps = 200 #number of steps of each run
         self.nEpchs = 8000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.05*self.nSteps) #number of steps of intrinsic policy

### 2023-06-18 plotFns.py
TriggerFound

@@ -159,6 +159,12 @@ def plotDat(fig, ax, p, rxFull, ryFull, phxFull, phyFull, rwdFull, valFn, alph):
             linewidth=0.5,
             alpha=alph/500,
             zorder=2)
+    ax[0].scatter(p.tgtOU[0], p.tgtOU[1],
+            s=10,
+            color='white',
+            edgecolor='k',
+            linewidth=0.5,
+            alpha=1)
     
     t = np.arange(np.size(rwdFull))*p.dt
     ax[1].plot(t, rwdFull,

### 2023-06-18 prmts.py

@@ -5,10 +5,10 @@ class params():
     def __init__(self):
         super().__init__()
         # self.nSteps = 1000 #number of steps of each run
-        self.nSteps = 200 #number of steps of each run
+        self.nSteps = 1000 #number of steps of each run
         self.nEpchs = 8000 #number of epochs
         self.a = 1.0 #radius of semi-circle
-        self.nSc = int(0.05*self.nSteps) #number of steps of intrinsic policy
+        self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
         
         # Pheromone parameters
         self.nMsh = 500 #number of points in grid
@@ -21,16 +21,16 @@ class params():
         self.dt = 1e-2 #time-step size
         self.l = 5e-3 #length travelled in 1 time-step
         self.vo = self.l/self.dt #effective speed of motion
-        self.nu = 10 #orientation relaxation rate (s^-1)
+        self.nu = 10. #orientation relaxation rate (s^-1)
         # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
-        self.diffCst = 0.1 #noise diffusion coefficient (s^-1)
+        self.diffCst = 1. #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
         
         # Reward/Learning parameters
         self.alpha = 0.7 #learning rate
-        self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
+        self.phiSt = np.pi/2 #reward goes as exp(-\phi/\phiSt)
         self.sigma = 0.1 #radius of region near target within which agent succeeds
         self.nPtn = 40 #number of division of \phi over which value function is define
         self.epsilon = 0.9 #constant for epsilon-greedy strategy
@@ -38,6 +38,7 @@ class params():
         # Initial conditions
         self.rInit = np.array([self.a + np.random.uniform(-self.pt, self.pt), 0.0]) #initial location of agent
         self.tgt = np.array([-self.a, 0.0]) #target location of agent
+        self.tgtOU = np.array([-self.a, 0.0]) #target location of agent
         self.thetInit = 0.5*np.pi #initial orientation of agent
         # self.rng = random.PRNGKey(1) #initial random number
         self.tgtPhi = 0.0 #target orientation from current location

### 2023-06-20 prmts.py
AlmostDone

@@ -6,7 +6,7 @@ class params():
         super().__init__()
         # self.nSteps = 1000 #number of steps of each run
         self.nSteps = 1000 #number of steps of each run
-        self.nEpchs = 8000 #number of epochs
+        self.nEpchs = 10000 #number of epochs
         self.a = 1.0 #radius of semi-circle
         self.nSc = int(0.1*self.nSteps) #number of steps of intrinsic policy
         
@@ -21,17 +21,17 @@ class params():
         self.dt = 1e-2 #time-step size
         self.l = 5e-3 #length travelled in 1 time-step
         self.vo = self.l/self.dt #effective speed of motion
-        self.nu = 10. #orientation relaxation rate (s^-1)
+        self.nu = 0.5 #orientation relaxation rate (s^-1)
         # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
-        self.diffCst = 1. #noise diffusion coefficient (s^-1)
+        self.diffCst = 0.1 #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
         
         # Reward/Learning parameters
         self.alpha = 0.7 #learning rate
-        self.phiSt = np.pi/2 #reward goes as exp(-\phi/\phiSt)
-        self.sigma = 0.1 #radius of region near target within which agent succeeds
+        self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
+        self.sigma = 0.3 #radius of region near target within which agent succeeds
         self.nPtn = 40 #number of division of \phi over which value function is define
         self.epsilon = 0.9 #constant for epsilon-greedy strategy
         

### 2023-06-21 prmts.py
CleanUpPartlyDone

@@ -22,11 +22,11 @@ class params():
         self.l = 5e-3 #length travelled in 1 time-step
         self.vo = self.l/self.dt #effective speed of motion
         self.nu = 0.5 #orientation relaxation rate (s^-1)
-        # self.diff = 50e-1 #noise diffusion coefficient (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
         self.diffCst = 0.1 #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
+        self.lNu = self.vo/self.nu #length-scale used for non-dimensionalization (s)
         
         # Reward/Learning parameters
         self.alpha = 0.7 #learning rate

### 2023-06-22 prmtsLin.py
Fig1CloseToReady

@@ -21,9 +21,9 @@ class params():
         self.dt = 1e-2 #time-step size
         self.l = 5e-3 #length travelled in 1 time-step
         self.vo = self.l/self.dt #effective speed of motion
-        self.nu = 50 #orientation relaxation rate (s^-1)
+        self.nu = 0.5 #orientation relaxation rate (s^-1)
         # self.diff = 10/(self.nSc*self.l) #noise diffusion coefficient (s^-1)
-        self.diffCst = 10 #noise diffusion coefficient (s^-1)
+        self.diffCst = 0.1 #noise diffusion coefficient (s^-1)
         self.diff = self.diffCst #dynamic diffusion coefficient (s^-1)
         self.pe = self.diff/self.nu #peclet number
         self.lNu = self.vo/self.nu #length-scale used for non-dimensionalization (s)
@@ -31,7 +31,7 @@ class params():
         # Reward/Learning parameters
         self.alpha = 0.7 #learning rate
         self.phiSt = np.pi/8 #reward goes as exp(-\phi/\phiSt)
-        self.sigma = 0.1 #radius of region near target within which agent succeeds
+        self.sigma = 0.3 #radius of region near target within which agent succeeds
         self.nPtn = 40 #number of division of \phi over which value function is define
         self.epsilon = 0.9 #constant for epsilon-greedy strategy
         

### 2023-06-23 LaTeX/draft.tex
TriggerFound
ErrorFound
rwdPlotsClear

@@
     \centering
-    \includegraphics[width=\textwidth]{./figs/schematic.pdf}\label{fig:schm1}
+    \includegraphics[width=\textwidth]{./figs/fig1.pdf}\label{fig:schm1}
     \caption{Schematic of setup}
@@
 can reach the target, then the probability that an agent will reach the target is given by $\P(||\r(s) - \r^*|| \leq \sigma)
-= 2\int_0^\sigma \mu \sqrt{\sigma^2-\zeta^2} \ \d \zeta/\psi(s^*-\zeta)$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
+= 2\int_0^\sigma \mu \sqrt{\sigma^2-\zeta^2} \ \d \zeta/\sqrt{\psi(s^*-\zeta)}$. We can now calculate the value function, $v_{\rm OU}(\r, \theta)$ as
+\textbf{***This cannot be done because trajectories have to be conditioned on the previous step if intersection at a particular location has to be accounted.***}
 \[

### 2023-06-23 LaTeX/draft.tex
DraftUpdates

@@
     \includegraphics[width=\textwidth]{./figs/fig1.pdf}\label{fig:schm1}
-    \caption{Schematic of setup}
+    \caption{$(a)$ Schematic of an agent described by its position $\r(t)$ and orientation $\ph (t)$
+    in an environment with pheromone already laid out at location $\x^*(s)$. Here $t$ is the time and
+    $s$ the arc-length along the trail with $\th(s)$ being the local tangent vector along the pheromone trail.
+    $(b)$ An agent traversing a familiar semi-circular pheromone trail (of radius $a$) from home at B to food source
+    at A. The task of the agent is to innovate from this semi-circular trail given uncertainities in its
+    memory of the target food source at A. $(c)$ The strategy used by the agent is split into two behaviors
+    represented as policies: $(i)$ intrinsic behavior to follow the already laid trail, $\pi_{\rm int} (\r, \phi)$;
+    $(ii)$ exploratory behaviour captured by an Ornstein-Uhlenbeck process in orientaiton, $\pi_{\rm OU}(\r, \phi^*; D)$.
+    $(d)$ Evolution of the value function, $v(s_c)$ learnt by the agent with epochs (details in SI.XXX) where $s_c$ is the
+    critical length at which innovation takes place. $(e)$ Evolution of the reward, $R(s)$ aquired by the agent 
+    after learning the value function. Sharp jumps in the magnitude are when the agent has reached the
+    target successfully. $(f)$ Comparison between theoretically calculated value function, $v(s_c)$ with
+    learnt value function by the agent for different values of non-dimensional numbers (see text for details),
+    the Peclet number, $\Pe$ and curvature, $\kappa$.
+    }
 \end{figure*}
